### 2017-07-03 paper.tex
Copied the README.md file across to a new .tex file, ahead of moving references to BiBTeX.

\documentclass{article}
\usepackage[a4paper]{geometry}
\begin{document}

\title{Ten Simple Rules for Teaching Programming}

\author{Greg Wilson and Neil C. C. Brown}

%1. King's College London / neil.c.c.brown@kcl.ac.uk
%2. Rangle.io / gvwilson@third-bit.com

\maketitle

\section*{Introduction}

Computing is a younger discipline than mathematics, physics, or biology,
and there have been correspondingly fewer studies of how best to teach it.
However,
there is a growing body of evidence about what works and doesn't in programming education.
This paper presents ten simple rules that should be the foundation of any course,
formal or informal.

FIXME: mention Repenning's boring/exciting X hard/easy framework from "Moving Beyond Syntax" paper.

\section{There's no geek gene} %(Patitsas et al)

Guzdial refers to the belief that some people are born programmers and others aren't
as [computing's most enduring and damaging myth][guzdial-geek].
Our first and most important rule is that this is wrong:
competence at programming is not innate,
but is rather a learned skill that can be acquired and improved with practice.

The most powerful evidence for this comes from [Patitsas et al][patitsas].
They examined grade distributions in introductory computing courses at a large university,
and found that only 5.8% were actually multi-modal.
More damningly,
they found that computer science faculty were more likely to see distributions as bimodal
if they thought those grades came from a programming class
than if they believed the grades came from some other kind of class,
and that those faculty were even more likely to see the distributions as bimodal
if they believed some students are innately predisposed to do well in computer science.

Beliefs such as this are known to have powerful effects on education outcomes [FIXME][fixme-outcomes].
If instructors believe that "some kids get it and some kids don't",
they will (consciously or unconsciously) invest less in those whom they put in the second category.
When combined with cultural stereotypes about who is and isn't a "natural programmer",
the downward spiral of under-achievement that results from differential attention
may be partly responsible for the gender imbalance in computing.

\section{Use live coding}

Rather than using slides,
instructors should write programs in front of their learners [FIXME][fixme-live-coding].
This is more effective for multiple reasons:

1.  It enables instructors to be more responsive to "what if?"
    questions. Where a slide deck is like a railway track, live coding
    allows instructors to go off road and follow their learners'
    interests.

2.  It facilitates unintended knowledge transfer: students learn more
    than the instructor consciously intends to teach by watching *how*
    instructors do things.

3.  It slows the instructor down: if they has to type in the program
    as they go along, they can only go twice as fast as their
    learners, rather than ten-fold faster as they could with slides.

4.  Learners get to how instructors diagnose and correct
    mistakes. Novices are going to spend most of their time doing
    this, but it's left out of most textbooks.

5.  Watching instructors make mistakes shows learners that it's all
    right to make mistakes of their own.  Most people model the
    behavior of their teachers: if the instructor isn't embarrassed
    about making and talking about mistakes, learners will be more
    comfortable doing so too.

Live coding does have some drawbacks, but with practice, these can be
avoided or worked around:

1.  Instructors can go too slowly, either because they are not good
    typists or by spending too much time looking at notes to try to
    remember what they meant to type.

2.  Typing in boilerplate code that is needed by the lesson, but not
    directly relevant to it (such as library import statements), can
    distract learners from the intended thrust of a lesson.  As
    [Willingham][willingham] says, "Memory is the residue of thought";
    if the instructor spends their time typing boilerplate, that may
    be all that learners take away.

\section{Have students make predictions}

- Being wrong stings!  Which is good.  Easy to brush it off if you haven't made prediction public.
- Research on how students misremembered physics demos afterwards.
- More engaging if you have to have to think about it and take an action.

\section{Use peer instruction}

No matter how good teachers are,
they can only say one thing at a time.
How then can they clear up learners' many different misconceptions
in a reasonable time?

The best method developed so far is called peer instruction.
Originally created by Eric Mazur at Harvard [FIXME][fixme-mazur],
it has been studied extensively in a wide variety of contexts,
including programming [Porter2013][porter] and [cutts][fixme-cutts].
In simplified form,
peer instruction proceeds in three phases

1. The instructor gives learners a brief introduction to the topic.
1. The instructor then gives learners a multiple choice question
   that probes for misconceptions rather than simple factual recall.
1. Learners then vote on the answer to the question.
   * If they all have the right answer, move on.
   * If they all have the same wrong answer,
     the instructor addresses that specific misconception.
   * If they have a mix of right and wrong answers,
     they are given several minutes to discuss those answers with one another
     in small groups (typically 2-4 students)
     and then reconvene and vote again.

Peer instruction is essentially
a way to provide one-to-one mentorship in a scalable way.
Group discussion significantly improves learners' understanding
because it forces them to clarify their thinking,
which can be enough to call out gaps in reasoning.
Re-polling the class then lets the instructor know if they can move on,
or if further explanation is necessary.
While it significantly outperforms lecture-based instruction in most situations,
it can be problematic if ability levels differ widely
(as they often do in introductory programming classes
because of varied prior experience).
The practice advocated in our next rule can be used to mitigate this.

\section{Use pair programming}

- Evens out differences in ability
- Both parties learn something
  - The weaker gets individual instruction
  - The stronger learns from teaching
- McDowell et al 2006

\section{Use worked examples with labelled subgoals}

- Lets you give students many different specific problems, while making clear the higher-level isomorphisms between the problems.
- Without subgoal labels, learners don't always see the spirals in a spiral curriculum.
- There is a complex result in the Morrison, Guzdial work about writing your own labels.  May be too intricate to include?

\section{Start with blocks}

- Evidence generally points to blocks being better than text.
- This can conflict with context though, and you may encounter resistance from learners (inauthenticity, etc)

\section{Stick to one language}

A principle that applies across all areas of education is that transference only comes with mastery [FIXME][fixme-transference]:
novices don't yet have a working mental model of the problem domain,
and so are unable to distinguish general principles from the specifics of particular problem instances.
While an experience programmer can,
for example,
take what they know about loops and function calls in one language
and re-use that understanding in a language with a different syntax or semantics,
a newcomer does not yet know which elements of their knowledge are central
and which are accidental.

Courses should therefore stick to one language until learners have progressed far enough with it
to be able to distinguish the forest from the trees.
Attempting to force transference too early---i.e.,
requiring them to switch from Python to JavaScript in order to do a web programming course
early in their education---will confuse learners and erode their confidence.

\section{Use authentic tasks}

[Guzdial et al][guzdial-media] found that having learners manipulate images, audio, and video
in their early programming assignments
increased retention in two senses:
learners remembered more of the material when re-tested after a delay,
and were more likely to stay in computing programs.
Their studies,
which have since been replicated [FIXME][fixme-media-comp-replication],
are consistent with findings from other areas of educational research [FIXME][fixme-authentic-tasks]:
learners find *authentic tasks* more engaging than abstracted examples.

FIXME: ITiCSE report says context makes no difference to task performance.

One caution is that context can inadvertently exclude some people while drawing others in.
For example,
many educators use computer games as a motivating example for programming classes,
but some learners may associate them with violence and racial or gender stereotypes.

\section{Don't just code}

Our final rule for teaching programming is that you don't have to program to do it.
Between syntax, semantics, algorithms, and design,
examples that seem small to instructors can easily overwhelm novices.
Breaking the problem down into smaller single-concept pieces can reduce the cognitive load to something manageable.

For example,
a growing number of educators are including Parsons Problems in their pedagogic repertoire [FIXME][fixme-parsons].
Rather than writing programs from scratch,
learners are given the lines of code they need to solve a problem,
but in jumbled order.
Re-ordering them to solve the problem correctly allows them to concentrate on mastering control flow
without having to devote mental energy to recalling syntax
or the specifics of library functions.

\section*{Conclusion}

Like anything involving human subjects,
studies of computing education must necessarily be hedged with qualifiers.
However,
we do know a great deal,
and are learning more each year.
Venues like [SIGCSE][sigcse],
[ITiCSE][iticse],
and [CSER][cser]
present a growing number of rigorous, insightful studies
with immediate practical application.
Future work may overturn or qualify some of our ten simple rules,
but they form a solid basis for any educational effort.

[fixme-authentic-tasks]: cite-value-of-authentic-tasks
[fixme-cutts]: cutts-ref-for-peer-instruction
[fixme-live-coding]: cite-live-coding
[fixme-mazur]: cite-mazur-peer-instruction
[fixme-media-comp-replication]: cite-media-computing-replication
[fixme-outcomes]: cite-belief-in-innate-ability-shaping-outcomes
[fixme-parsons]: cite-parsons-problems
[fixme-transference]: cite-on-general-knowledge-transfer
[guzdial-geek]: https://cacm.acm.org/blogs/blog-cacm/189498-top-10-myths-about-teaching-computer-science/fulltext
[guzdial-media]: FIXME
[patitsas]: http://www.cs.toronto.edu/~sme/papers/2016/icer\_2016\_bimodal.pdf
[porter]: porter-et-al-peer-instruction
[sigcse]: FIXME
[willingham]: https://www.amazon.com/Why-Dont-Students-Like-School/dp/047059196X/

\end{document}

### 2017-07-03 paper.tex
Moved the first few references across to BiBTeX

@@
 \usepackage[a4paper]{geometry}
+\usepackage{natbib}
+\usepackage{hyperref}
 \begin{document}
@@
 Guzdial refers to the belief that some people are born programmers and others aren't
-as [computing's most enduring and damaging myth][guzdial-geek].
+as ``computing's most enduring and damaging myth'' \citep{guzdial-myths}.
 Our first and most important rule is that this is wrong:
@@
 
-The most powerful evidence for this comes from [Patitsas et al][patitsas].
+The most powerful evidence for this comes from \citet{patitsas-bimodal}.
 They examined grade distributions in introductory computing courses at a large university,
-and found that only 5.8% were actually multi-modal.
+and found that only 5.8\% were actually multi-modal.
 More damningly,
@@
     distract learners from the intended thrust of a lesson.  As
-    [Willingham][willingham] says, "Memory is the residue of thought";
+    \citet{willingham-book} says, "Memory is the residue of thought";
     if the instructor spends their time typing boilerplate, that may
@@
 [fixme-transference]: cite-on-general-knowledge-transfer
-[guzdial-geek]: https://cacm.acm.org/blogs/blog-cacm/189498-top-10-myths-about-teaching-computer-science/fulltext
 [guzdial-media]: FIXME
-[patitsas]: http://www.cs.toronto.edu/~sme/papers/2016/icer\_2016\_bimodal.pdf
 [porter]: porter-et-al-peer-instruction
 [sigcse]: FIXME
-[willingham]: https://www.amazon.com/Why-Dont-Students-Like-School/dp/047059196X/
+
+\bibliographystyle{abbrvnat}
+\bibliography{refs}
 

### 2017-07-03 paper.tex
Filled in a couple of sections and added some references.

@@
 
-1.  It enables instructors to be more responsive to "what if?"
+\begin{enumerate}
+\item It enables instructors to be more responsive to "what if?"
     questions. Where a slide deck is like a railway track, live coding
@@
 
-2.  It facilitates unintended knowledge transfer: students learn more
+\item  It facilitates unintended knowledge transfer: students learn more
     than the instructor consciously intends to teach by watching *how*
@@
 
-3.  It slows the instructor down: if they has to type in the program
+\item  It slows the instructor down: if they has to type in the program
     as they go along, they can only go twice as fast as their
@@
 
-4.  Learners get to how instructors diagnose and correct
+\item  Learners get to how instructors diagnose and correct
     mistakes. Novices are going to spend most of their time doing
@@
 
-5.  Watching instructors make mistakes shows learners that it's all
-    right to make mistakes of their own.  Most people model the
+\item  Watching instructors make mistakes shows learners that it's all
+    right to make mistakes of their own~\citep{barker-live-coding}.  Most people model the
     behavior of their teachers: if the instructor isn't embarrassed
@@
     comfortable doing so too.
+\end{enumerate}
 
@@
 
-1.  Instructors can go too slowly, either because they are not good
+\begin{enumerate}
+\item  Instructors can go too slowly, either because they are not good
     typists or by spending too much time looking at notes to try to
@@
 
-2.  Typing in boilerplate code that is needed by the lesson, but not
+\item  Typing in boilerplate code that is needed by the lesson, but not
     directly relevant to it (such as library import statements), can
@@
     be all that learners take away.
+\end{enumerate}
 
@@
 
-- Being wrong stings!  Which is good.  Easy to brush it off if you haven't made prediction public.
-- Research on how students misremembered physics demos afterwards.
-- More engaging if you have to have to think about it and take an action.
+TODO I wonder if this section should be merged with peer instruction?  It overlaps a lot.
+
+Live coding is a form of demonstration.
+The instructor shows how to write a program, and will usually run the program once or more during its development to show what it does.
+Research from peer instruction in physics education shows that learners who observe a demonstration do not learn better than those who did not see the demonstration~\citep{crouch-demo}, and in fact many learners misremember the outcome of demonstrations afterwards~\citep{miller-demo}.
+
+The key to making demonstrations more effective is to make learners predict the outcome of the demonstration before performing it.
+Crucially, their prediction should be in some way recorded or public.
+This might be by a show of hands, or by holding up a cue card (A/B/C/D), or by talking to their neighbour.
+We believe it is the sting of being publicly wrong that enhances the learning.
+Naturally, the instructor should take care not to punish or criticise those students who predicted wrongly.
 
@@
 The best method developed so far is called peer instruction.
-Originally created by Eric Mazur at Harvard [FIXME][fixme-mazur],
+Originally created by Eric Mazur at Harvard~\citep{mazur-peer-instruction},
 it has been studied extensively in a wide variety of contexts,
@@
 
-- Lets you give students many different specific problems, while making clear the higher-level isomorphisms between the problems.
-- Without subgoal labels, learners don't always see the spirals in a spiral curriculum.
-- There is a complex result in the Morrison, Guzdial work about writing your own labels.  May be too intricate to include?
+Learning to program involves learning the syntax and semantics of a programming language, but also involves learning how to construct programs.
+A good way to guide students through constructing programs is the use of worked examples: step-by-step guides showing how to solve an existing problem.
+
+A good idea when teaching is to provide many similar examples for learners to practice on.
+But since learners are novices, they may not see the similarity between examples (such as finding the highest rainfall from a list of numbers, versus finding the first surname alphabetically from a list of names).
+Subgoal labels break the worked examples down into more manageable chunks, and they also provide a structure which students can apply to future tasks that they attempt themselves.
+\citet{morrison-subgoals} has shown that students perform better with subgoal labels than without.
+
+TODO should we give a short code snippet here with some subgoal labels to illustrate?
 
@@
 [fixme-cutts]: cutts-ref-for-peer-instruction
-[fixme-live-coding]: cite-live-coding
-[fixme-mazur]: cite-mazur-peer-instruction
 [fixme-media-comp-replication]: cite-media-computing-replication

### 2017-07-04 paper.tex
Added a few more references, and swapped out the blocks principle for one about novices-vs-experts.

@@
 
-Computing is a younger discipline than mathematics, physics, or biology,
-and there have been correspondingly fewer studies of how best to teach it.
-However,
-there is a growing body of evidence about what works and doesn't in programming education.
-This paper presents ten simple rules that should be the foundation of any course,
+Research from educational psychology suggests that teaching and learning is a subject-specific
+activity~\citep{mayer-subject}: learning programming has a different set of challenges and techniques
+than learning physics or learning to read and write.  Computing is a younger discipline than
+mathematics, physics, or biology, and there have been correspondingly fewer studies of how best to teach it.
+However, there is a growing body of evidence about what works and doesn't in programming education.
+This paper presents ten simple rules that should be the foundation of any teaching, whether
 formal or informal.
@@
 
+\textbf{Note: I have several reservations about the Rubin paper; it doesn't really show much difference for live coding, and it uses debunked learning styles, although that's not too important to the outcome.
+But it turns out there's very few papers on live coding.
+The Barker paper is generally better, but only mentions live coding tangentially.}
+
 Rather than using slides,
-instructors should write programs in front of their learners [FIXME][fixme-live-coding].
+instructors should write programs in front of their learners~\cite{rubin-live-coding}.
 This is more effective for multiple reasons:
@@
 
-TODO I wonder if this section should be merged with peer instruction?  It overlaps a lot.
+\textbf{TODO I wonder if this section should be merged with peer instruction?  It overlaps a lot.}
 
@@
 Subgoal labels break the worked examples down into more manageable chunks, and they also provide a structure which students can apply to future tasks that they attempt themselves.
-\citet{morrison-subgoals} has shown that students perform better with subgoal labels than without.
+\citet{morrison-subgoals,morrison-parsons} have shown that students perform better with subgoal labels than without.
 
@@
 
-\section{Start with blocks}
+\section{Novices are not experts}
 
-- Evidence generally points to blocks being better than text.
-- This can conflict with context though, and you may encounter resistance from learners (inauthenticity, etc)
+This principle sounds tautological, but it is easily forgotten.
+Novices program differently than experts, and need different approaches or tools.
+If you ask a professional programmer to iterate over a list of integers and produce the average, they can write the code within seconds, using stored knowledge of the exact pattern required.
+A novice will approach this problem totally differently.
+They need to remember the syntax for the different parts, they need to know how to iterate over a list, how to use an accumulator variable and so on.
+
+Novices may need to spend time thinking about an algorithm on paper (something expert programmers rarely need, as they have usually memorised most common algorithmic patterns).
+They may need to construct examples in guided steps.
+They may struggle to debug.
+Debugging usually involves contrasting what is happening to what should be happening, but a novice's grasp on what should be happening is usually fragile.
+
+Novices do not become professionals simply by doing what professionals do at a slower pace.
+We do not teach reading by taking a classic novel and simply proceeding more slowly.
+We teach by using shorter books with simpler words and larger print.
+So in programming, we must take care to use small, self-contained tasks at a level suitable for novices, with tools that suit their needs.
 
@@
 
-FIXME: ITiCSE report says context makes no difference to task performance.
+A classic question in computing (and mathematics) education is whether problems are better 
+with context (e.g. find the highest student grade) or without (e.g. find the maximum of the list of numbers).
+\citet{bouvier-context} studied this with a multi-university study and found no difference between
+the two.  They suggest that since it makes no difference, other considerations (such as motivation)
+should be primary.
 
@@
 For example,
-a growing number of educators are including Parsons Problems in their pedagogic repertoire [FIXME][fixme-parsons].
+a growing number of educators are including Parson's Problems in their pedagogic repertoire~\citep{parsons,morrison-parsons}.
 Rather than writing programs from scratch,
@@
 [fixme-outcomes]: cite-belief-in-innate-ability-shaping-outcomes
-[fixme-parsons]: cite-parsons-problems
 [fixme-transference]: cite-on-general-knowledge-transfer

### 2017-07-04 paper.tex
A few small tidy-ups, adjusting from Markdown to LaTeX

@@
 
-Guzdial refers to the belief that some people are born programmers and others aren't
-as ``computing's most enduring and damaging myth'' \citep{guzdial-myths}.
+\citet{guzdial-myths} refers to the belief that some people are born programmers and others aren't
+as ``computing's most enduring and damaging myth.''
 Our first and most important rule is that this is wrong:
@@
 Beliefs such as this are known to have powerful effects on education outcomes [FIXME][fixme-outcomes].
-If instructors believe that "some kids get it and some kids don't",
+If instructors believe that ``some kids get it and some kids don't'',
 they will (consciously or unconsciously) invest less in those whom they put in the second category.
@@
 \begin{enumerate}
-\item It enables instructors to be more responsive to "what if?"
+\item It enables instructors to be more responsive to ``what if?''
     questions. Where a slide deck is like a railway track, live coding
@@
 \item  It facilitates unintended knowledge transfer: students learn more
-    than the instructor consciously intends to teach by watching *how*
+    than the instructor consciously intends to teach by watching \textit{how}
     instructors do things.
@@
     distract learners from the intended thrust of a lesson.  As
-    \citet{willingham-book} says, "Memory is the residue of thought";
+    \citet{willingham-book} says, ``Memory is the residue of thought'';
     if the instructor spends their time typing boilerplate, that may
@@
 
-1. The instructor gives learners a brief introduction to the topic.
-1. The instructor then gives learners a multiple choice question
+\begin{enumerate}
+\item The instructor gives learners a brief introduction to the topic.
+\item The instructor then gives learners a multiple choice question
    that probes for misconceptions rather than simple factual recall.
-1. Learners then vote on the answer to the question.
-   * If they all have the right answer, move on.
-   * If they all have the same wrong answer,
+\item Learners then vote on the answer to the question.
+  \begin{itemize}
+   \item If they all have the right answer, move on.
+   \item If they all have the same wrong answer,
      the instructor addresses that specific misconception.
-   * If they have a mix of right and wrong answers,
+   \item If they have a mix of right and wrong answers,
      they are given several minutes to discuss those answers with one another
@@
      and then reconvene and vote again.
+\end{itemize}
+\end{enumerate}
 
@@
 For example,
-a growing number of educators are including Parson's Problems in their pedagogic repertoire~\citep{parsons,morrison-parsons}.
+a growing number of educators are including Parsons Problems in their pedagogic repertoire~\citep{parsons,morrison-parsons}.
 Rather than writing programs from scratch,

### 2017-07-04 paper.tex
Filling in

@@
 \usepackage{hyperref}
-\begin{document}
 
 \title{Ten Simple Rules for Teaching Programming}
+\author{Neil C. C. Brown\footnote{King's College London / neil.c.c.brown@kcl.ac.uk} and Greg Wilson\footnote{Rangle.io / gvwilson@third-bit.com}}
 
-\author{Greg Wilson and Neil C. C. Brown}
-
-%1. King's College London / neil.c.c.brown@kcl.ac.uk
-%2. Rangle.io / gvwilson@third-bit.com
+\begin{document}
 
@@
 
-Research from educational psychology suggests that teaching and learning is a subject-specific
-activity~\citep{mayer-subject}: learning programming has a different set of challenges and techniques
+Research from educational psychology suggests that teaching and learning are subject-specific
+activities~\citep{mayer-subject}: learning programming has a different set of challenges and techniques
 than learning physics or learning to read and write.  Computing is a younger discipline than
-mathematics, physics, or biology, and there have been correspondingly fewer studies of how best to teach it.
-However, there is a growing body of evidence about what works and doesn't in programming education.
+mathematics, physics, or biology, and while there have been correspondingly fewer studies of how best to teach it,
+there is a growing body of evidence about what works and doesn't.
 This paper presents ten simple rules that should be the foundation of any teaching, whether
@@
 
-FIXME: mention Repenning's boring/exciting X hard/easy framework from "Moving Beyond Syntax" paper.
+% FIXME: mention Repenning's boring/exciting X hard/easy framework from "Moving Beyond Syntax" paper.
 
-\section{There's no geek gene} %(Patitsas et al)
+\section{There is no geek gene}\label{gene}
 
@@
 
-Beliefs such as this are known to have powerful effects on education outcomes [FIXME][fixme-outcomes].
+Beliefs such as this are known to have powerful effects on education outcomes
+\cite{alvidrez-weinstein,brophy,jussim-eccles}.
 If instructors believe that ``some kids get it and some kids don't'',
@@
 
-\section{Use live coding}
+\section{Use peer instruction}\label{peer-instruction}
 
-\textbf{Note: I have several reservations about the Rubin paper; it doesn't really show much difference for live coding, and it uses debunked learning styles, although that's not too important to the outcome.
-But it turns out there's very few papers on live coding.
-The Barker paper is generally better, but only mentions live coding tangentially.}
+No matter how good teachers are,
+they can only say one thing at a time.
+How then can they clear up learners' many different misconceptions
+in a reasonable time?
+
+The best method developed so far is called peer instruction.
+Originally created by Eric Mazur at Harvard~\citep{mazur-peer-instruction},
+it has been studied extensively in a wide variety of contexts,
+including programming \cite{porter,cutts}
+In simplified form,
+peer instruction proceeds in three phases:
+
+\begin{enumerate}
+
+\item The instructor gives learners a brief introduction to the topic.
+
+\item The instructor then gives learners a multiple choice question
+   that probes for misconceptions rather than simple factual recall.
+
+\item Learners then vote on the answer to the question.
+
+  \begin{itemize}
+  \item If they all have the right answer, move on.
+  \item If they all have the same wrong answer,
+    the instructor addresses that specific misconception.
+  \item If they have a mix of right and wrong answers,
+    they are given several minutes to discuss those answers with one another
+    in small groups (typically 2-4 students)
+    and then reconvene and vote again.
+  \end{itemize}
+
+\end{enumerate}
+
+Peer instruction is essentially
+a way to provide one-to-one mentorship in a scalable way.
+Group discussion significantly improves learners' understanding
+because it forces them to clarify their thinking,
+which can be enough to call out gaps in reasoning.
+Re-polling the class then lets the instructor know if they can move on,
+or if further explanation is necessary.
+While it significantly outperforms lecture-based instruction in most situations,
+it can be problematic if ability levels differ widely
+(as they often do in introductory programming classes
+because of varied prior experience).
+Pair programming (Rule~\ref{pair-programming}) can be used to mitigate this.
+
+\section{Use live coding}\label{live-coding}
 
 Rather than using slides,
-instructors should write programs in front of their learners~\cite{rubin-live-coding}.
+create programs in front of their learners~\cite{rubin-live-coding}.
 This is more effective for multiple reasons:
@@
 \begin{enumerate}
+
 \item It enables instructors to be more responsive to ``what if?''
@@
 \item Watching instructors make mistakes shows learners that it's all
-    right to make mistakes of their own~\citep{barker-live-coding}.  Most people model the
-    behavior of their teachers: if the instructor isn't embarrassed
-    about making and talking about mistakes, learners will be more
-    comfortable doing so too.
+  right to make mistakes of their own~\citep{barker-live-coding}.
+  Most people model the behavior of their teachers: if the instructor
+  isn't embarrassed about making and talking about mistakes, learners
+  will be more comfortable doing so too.
+
 \end{enumerate}
@@
 \begin{enumerate}
+
 \item Instructors can go too slowly, either because they are not good
@@
   \citet{willingham-book} says, ``Memory is the residue of thought'';
-    if the instructor spends their time typing boilerplate, that may
-    be all that learners take away.
+  if the instructor spends their time typing boilerplate, that may be
+  all that learners take away.
+
 \end{enumerate}
 
-\section{Have students make predictions}
+\section{Have students make predictions}\label{predictions}
 
-\textbf{TODO I wonder if this section should be merged with peer instruction?  It overlaps a lot.}
+When instructors are using live coding,
+they usually run the program several times during its development to show what it does.
+Research from peer instruction in physics education shows that
+learners who observe a demonstration do not learn better than
+those who did not see the demonstration~\citep{crouch-demo},
+and in fact many learners misremember the outcome of demonstrations afterwards~\citep{miller-demo}.
 
-Live coding is a form of demonstration.
-The instructor shows how to write a program, and will usually run the program once or more during its development to show what it does.
-Research from peer instruction in physics education shows that learners who observe a demonstration do not learn better than those who did not see the demonstration~\citep{crouch-demo}, and in fact many learners misremember the outcome of demonstrations afterwards~\citep{miller-demo}.
+The key to making demonstrations more effective is
+to make learners predict the outcome of the demonstration before performing it.
+Crucially, their prediction should be in some way recorded or public,
+e.g. by a show of hands,
+by holding up a cue card (A/B/C/D),
+or by talking to their neighbour.
+We speculate that the sting of possibly being publicly wrong leads learners to pay more attention
+and to reflect on what they are learning;
+regardless of whether this hypothesis is true,
+instructors should be careful not to punish or criticise students who predicted wrongly,
+but rather to use those incorrect predictions as a spur to further exploration.
 
-The key to making demonstrations more effective is to make learners predict the outcome of the demonstration before performing it.
-Crucially, their prediction should be in some way recorded or public.
-This might be by a show of hands, or by holding up a cue card (A/B/C/D), or by talking to their neighbour.
-We believe it is the sting of being publicly wrong that enhances the learning.
-Naturally, the instructor should take care not to punish or criticise those students who predicted wrongly.
+\section{Use pair programming}\label{pair-programming}
 
-\section{Use peer instruction}
+Pair programming is a software development practice
+in which two programmers share one computer.
+One person (called the driver) does the typing,
+while the other (called the navigator) offers comments and suggestions.
+The two switch roles several times per hour.
+Pair programming is a good practice in real-life programming \cite{hannay},
+and also a good way to teach \cite{mcdowell}.
+Partners can not only help each other out during practical exercises,
+but can also clarify each other's misconceptions when the solution is presented.
 
-No matter how good teachers are,
-they can only say one thing at a time.
-How then can they clear up learners' many different misconceptions
-in a reasonable time?
+Both parties involved in pair programming learn while doing it.
+The weaker gets individual instruction from the stronger,
+while the stronger learns by explaining,
+and by being forced to reconsider things which they may not have thought about in a while.
+When pair programming is used it is important to put everyone in pairs,
+not just the learners who may be struggling,
+so that no one feels singled out.
+It's also important to have people switch roles within each pair three or four times per hour,
+so that the stronger personality in each pair does not dominate the session.
 
-The best method developed so far is called peer instruction.
-Originally created by Eric Mazur at Harvard~\citep{mazur-peer-instruction},
-it has been studied extensively in a wide variety of contexts,
-including programming [Porter2013][porter] and [cutts][fixme-cutts].
-In simplified form,
-peer instruction proceeds in three phases
+\section{Use worked examples with labelled subgoals}\label{worked-examples}
 
-\begin{enumerate}
-\item The instructor gives learners a brief introduction to the topic.
-\item The instructor then gives learners a multiple choice question
-   that probes for misconceptions rather than simple factual recall.
-\item Learners then vote on the answer to the question.
-  \begin{itemize}
-   \item If they all have the right answer, move on.
-   \item If they all have the same wrong answer,
-     the instructor addresses that specific misconception.
-   \item If they have a mix of right and wrong answers,
-     they are given several minutes to discuss those answers with one another
-     in small groups (typically 2-4 students)
-     and then reconvene and vote again.
-\end{itemize}
-\end{enumerate}
+Learning to program involves learning the syntax and semantics of a programming language,
+but also involves learning how to construct programs.
+A good way to guide students through constructing programs is the use of worked examples:
+step-by-step guides showing how to solve an existing problem.
 
-Peer instruction is essentially
-a way to provide one-to-one mentorship in a scalable way.
-Group discussion significantly improves learners' understanding
-because it forces them to clarify their thinking,
-which can be enough to call out gaps in reasoning.
-Re-polling the class then lets the instructor know if they can move on,
-or if further explanation is necessary.
-While it significantly outperforms lecture-based instruction in most situations,
-it can be problematic if ability levels differ widely
-(as they often do in introductory programming classes
-because of varied prior experience).
-The practice advocated in our next rule can be used to mitigate this.
+A good idea when teaching is
+to provide many similar examples for learners to practice on.
+But since learners are novices,
+they may not see the similarity between examples:
+finding the highest rainfall from a list of numbers
+finding the first surname alphabetically from a list of names
+may seem like quite different problems to them.
 
-\section{Use pair programming}
+\citet{morrison-subgoals,morrison-parsons} have shown that students perform better
+when worked examples are broken down into steps (or subgoals)
+which are given names (or labels).
+Subgoal labels provide a structure which students can apply to future tasks that they attempt themselves,
+and allow them to communicate with their peers and instructors more efficiently.
 
-- Evens out differences in ability
-- Both parties learn something
-  - The weaker gets individual instruction
-  - The stronger learns from teaching
-- McDowell et al 2006
+% FIXME give a short code snippet here with some subgoal labels to illustrate - yes please
 
-\section{Use worked examples with labelled subgoals}
+\section{Stick to one language}\label{one-language}
 
-Learning to program involves learning the syntax and semantics of a programming language, but also involves learning how to construct programs.
-A good way to guide students through constructing programs is the use of worked examples: step-by-step guides showing how to solve an existing problem.
-
-A good idea when teaching is to provide many similar examples for learners to practice on.
-But since learners are novices, they may not see the similarity between examples (such as finding the highest rainfall from a list of numbers, versus finding the first surname alphabetically from a list of names).
-Subgoal labels break the worked examples down into more manageable chunks, and they also provide a structure which students can apply to future tasks that they attempt themselves.
-\citet{morrison-subgoals,morrison-parsons} have shown that students perform better with subgoal labels than without.
-
-TODO should we give a short code snippet here with some subgoal labels to illustrate?
-
-\section{Novices are not experts}
-
-This principle sounds tautological, but it is easily forgotten.
-Novices program differently than experts, and need different approaches or tools.
-If you ask a professional programmer to iterate over a list of integers and produce the average, they can write the code within seconds, using stored knowledge of the exact pattern required.
-A novice will approach this problem totally differently.
-They need to remember the syntax for the different parts, they need to know how to iterate over a list, how to use an accumulator variable and so on.
-
-Novices may need to spend time thinking about an algorithm on paper (something expert programmers rarely need, as they have usually memorised most common algorithmic patterns).
-They may need to construct examples in guided steps.
-They may struggle to debug.
-Debugging usually involves contrasting what is happening to what should be happening, but a novice's grasp on what should be happening is usually fragile.
-
-Novices do not become professionals simply by doing what professionals do at a slower pace.
-We do not teach reading by taking a classic novel and simply proceeding more slowly.
-We teach by using shorter books with simpler words and larger print.
-So in programming, we must take care to use small, self-contained tasks at a level suitable for novices, with tools that suit their needs.
-
-\section{Stick to one language}
-
-A principle that applies across all areas of education is that transference only comes with mastery [FIXME][fixme-transference]:
-novices don't yet have a working mental model of the problem domain,
-and so are unable to distinguish general principles from the specifics of particular problem instances.
+A principle that applies across all areas of education is that transference only comes with mastery \cite{gick-holyoak}.
+Courses should therefore stick to one language until learners have progressed far enough with it
+to be able to distinguish the forest from the trees.
 While an experience programmer can,
@@
 and which are accidental.
-
-Courses should therefore stick to one language until learners have progressed far enough with it
-to be able to distinguish the forest from the trees.
 Attempting to force transference too early---i.e.,
@@
 
-\section{Use authentic tasks}
+\section{Use authentic tasks}\label{authentic-tasks}
 
-[Guzdial et al][guzdial-media] found that having learners manipulate images, audio, and video
+Guzdial et al.\ found that having learners manipulate images, audio, and video
 in their early programming assignments
@@
 learners remembered more of the material when re-tested after a delay,
-and were more likely to stay in computing programs.
-Their studies,
-which have since been replicated [FIXME][fixme-media-comp-replication],
-are consistent with findings from other areas of educational research [FIXME][fixme-authentic-tasks]:
-learners find *authentic tasks* more engaging than abstracted examples.
+and were more likely to stay in computing programs \cite{guzdial-media}.
+This is a particular instance of a larger observation:
+learners find authentic tasks more engaging than abstracted examples.
 
@@
 
-\section{Don't just code}
+\section{Don't just code}\label{not-just-code}
 
@@
 
+\section{Novices are not experts}\label{not-experts}
+
+This principle sounds tautological, but it is easily forgotten.
+Novices program differently than experts,
+and need different approaches or tools.
+If you ask a professional programmer to iterate over a list of integers and produce the average,
+they can write the code within seconds,
+using stored knowledge of the exact pattern required.
+A novice will approach this problem totally differently:
+they need to remember the syntax for the different parts,
+they need to know how to iterate over a list,
+how to use an accumulator variable,
+and so on.
+
+Novices may need to spend time thinking about an algorithm on paper
+(something expert programmers rarely need,
+as they have usually memorised most common algorithmic patterns).
+They may need to construct examples in guided steps.
+They may struggle to debug.
+Debugging usually involves contrasting what is happening to what should be happening,
+but a novice's grasp on what should be happening is usually fragile.
+
+Novices do not become professionals simply by doing what professionals do at a slower pace.
+We do not teach reading by taking a classic novel and simply proceeding more slowly.
+We teach by using shorter books with simpler words and larger print.
+So in programming,
+we must take care to use small, self-contained tasks at a level suitable for novices,
+with tools that suit their needs.
+
 \section*{Conclusion}
@@
 and are learning more each year.
-Venues like [SIGCSE][sigcse],
-[ITiCSE][iticse],
-and [CSER][cser]
+Venues like SIGCSE (\url{http://sigcse.org/}) and ITiCSE (\url{http://iticse.acm.org/})
 present a growing number of rigorous, insightful studies
@@
 
-[fixme-authentic-tasks]: cite-value-of-authentic-tasks
-[fixme-cutts]: cutts-ref-for-peer-instruction
-[fixme-media-comp-replication]: cite-media-computing-replication
-[fixme-outcomes]: cite-belief-in-innate-ability-shaping-outcomes
-[fixme-transference]: cite-on-general-knowledge-transfer
-[guzdial-media]: FIXME
-[porter]: porter-et-al-peer-instruction
-[sigcse]: FIXME
-
 \bibliographystyle{abbrvnat}

### 2017-07-07 paper.tex
Using paper.* for all names
Added a figure, plus various small fixes after a read-through.

@@
 there is a growing body of evidence about what works and doesn't.
-This paper presents ten simple rules that should be the foundation of any teaching, whether
+This paper presents ten simple rules that should be the foundation of any teaching of programming, whether
 formal or informal.
@@
 \citet{guzdial-myths} refers to the belief that some people are born programmers and others aren't
-as ``computing's most enduring and damaging myth.''
+as ``computing's most enduring and damaging myth.''  This is often ``confirmed'' by looking at university grade distributions, which are commonly held to be bimodal: a low-scoring hump of those that will never get it, and a high-scoring hump of those that have the right stuff.
 Our first and most important rule is that this is wrong:
@@
 Beliefs such as this are known to have powerful effects on education outcomes
-\cite{alvidrez-weinstein,brophy,jussim-eccles}.
+\citep{alvidrez-weinstein,brophy,jussim-eccles}.
 If instructors believe that ``some kids get it and some kids don't'',
@@
 
-No matter how good teachers are,
-they can only say one thing at a time.
-How then can they clear up learners' many different misconceptions
-in a reasonable time?
+One-on-one tutoring is perhaps the ideal form of teaching: all of a teacher's attention can be focused on one student, and they can completely customise their teaching for their single student, and tailor individual feedback and corrections based on two-way dialogue with that student.
+In realistic settings however, one teacher must teach many students at once, whether that is a handful, tens or even hundreds in large lectures.
+In larger settings, how can teachers possibly hope to clear up learners' many different misconceptions in a reasonable time?
 
-The best method developed so far is called peer instruction.
+The best method developed so far for larger-scale classrooms is called Peer Instruction.
 Originally created by Eric Mazur at Harvard~\citep{mazur-peer-instruction},
 it has been studied extensively in a wide variety of contexts,
-including programming \cite{porter,cutts}
+including programming \citep{cutts,porter}.
 In simplified form,
@@
 \item The instructor then gives learners a multiple choice question
-   that probes for misconceptions rather than simple factual recall.
+   that probes for misconceptions rather than simple factual recall.  A programming example is given in figure \ref{fig:mcq} which relates to integer comparison and loops.
 
@@
 
+\begin{figure}\sffamily
+\begin{verbatim}
+for (int i = 1; i < 10; i++)
+{
+     if (i < 3 || i > 8)
+     {
+          System.out.println("Yes");
+     }
+}
+\end{verbatim}
+How many times will the above code print out the word Yes?\\
+a) 10\\
+b) 9\\
+c) 4\\
+d) 3
+\label{fig:mcq}
+\caption{An example multiple choice question, probing learners' understanding of loops and integer comparisons.}
+\end{figure}
+
 \section{Use live coding}\label{live-coding}
@@
 Rather than using slides,
-create programs in front of their learners~\cite{rubin-live-coding}.
+instructors should create programs in front of their learners~\citep{rubin-live-coding}.
 This is more effective for multiple reasons:
@@
 \item It enables instructors to be more responsive to ``what if?''
-  questions. Where a slide deck is like a railway track, live coding
+  questions. Where a slide deck is like a highway, live coding
   allows instructors to go off road and follow their learners'
@@
   than the instructor consciously intends to teach by watching \textit{how}
-  instructors do things.
+  instructors do things.  The extra knowledge may be high-level (e.g. whether a program is written top-down or bottom-up) or fairly low-level (e.g. using keyboard shortcuts).
 
-\item It slows the instructor down: if they has to type in the program
+\item It slows the instructor down: if the instructor has to type in the program
   as they go along, they can only go twice as fast as their
@@
 
-\item Learners get to how instructors diagnose and correct
+\item Learners get to see how instructors diagnose and correct
   mistakes. Novices are going to spend most of their time doing
@@
 
-\item Watching instructors make mistakes shows learners that it's all
-  right to make mistakes of their own~\citep{barker-live-coding}.
+\item Watching instructors make mistakes shows learners that it's alright to make mistakes of their own~\citep{barker-live-coding}.
   Most people model the behavior of their teachers: if the instructor
@@
 
-\item Typing in boilerplate code that is needed by the lesson, but not
+\item Instructors can spend too much time typing in boilerplate code that is needed by the lesson, but not
   directly relevant to it (such as library import statements), can
@@
   if the instructor spends their time typing boilerplate, that may be
-  all that learners take away.
+  all that learners take away.  This can be avoided by starting with a partial skeleton that includes the boilerplate, or having it on hand to copy and paste at the appropriate juncture.
 
@@
 Research from peer instruction in physics education shows that
-learners who observe a demonstration do not learn better than
+learners who observe a demonstration \textit{do not} learn better than
 those who did not see the demonstration~\citep{crouch-demo},
-and in fact many learners misremember the outcome of demonstrations afterwards~\citep{miller-demo}.
+and in fact many learners misremember the outcome of demonstrations afterwards~\citep{miller-demo}.  That is, demonstrations can be useless or actively harmful to learning.
 
@@
 or by talking to their neighbour.
-We speculate that the sting of possibly being publicly wrong leads learners to pay more attention
+We speculate that the sting of being publicly wrong leads learners to pay more attention
 and to reflect on what they are learning;
@@
 The two switch roles several times per hour.
-Pair programming is a good practice in real-life programming \cite{hannay},
-and also a good way to teach \cite{mcdowell}.
+Pair programming is a good practice in real-life programming \citep{hannay},
+and also a good way to teach \citep{mcdowell}.
 Partners can not only help each other out during practical exercises,
@@
 they may not see the similarity between examples:
-finding the highest rainfall from a list of numbers
+finding the highest rainfall from a list of numbers and
 finding the first surname alphabetically from a list of names
-may seem like quite different problems to them.
+may seem like quite different problems to them, even though more advanced programmers would recognise them as isomorphic.
 
@@
 and allow them to communicate with their peers and instructors more efficiently.
+TODO give an example in a figure.
 
@@
 
-A principle that applies across all areas of education is that transference only comes with mastery \cite{gick-holyoak}.
+A principle that applies across all areas of education is that transference only comes with mastery \citep{gick-holyoak}.
 Courses should therefore stick to one language until learners have progressed far enough with it
 to be able to distinguish the forest from the trees.
-While an experience programmer can,
+While an experienced programmer can,
 for example,
@@
 and which are accidental.
-Attempting to force transference too early---i.e.,
+Attempting to force transference too early---e.g.,
 requiring them to switch from Python to JavaScript in order to do a web programming course
@@
 learners remembered more of the material when re-tested after a delay,
-and were more likely to stay in computing programs \cite{guzdial-media}.
+and were more likely to stay in computing programs \citep{guzdial-media}.
 This is a particular instance of a larger observation:
@@
 many educators use computer games as a motivating example for programming classes,
-but some learners may associate them with violence and racial or gender stereotypes.
-
-\section{Don't just code}\label{not-just-code}
-
-Our final rule for teaching programming is that you don't have to program to do it.
-Between syntax, semantics, algorithms, and design,
-examples that seem small to instructors can easily overwhelm novices.
-Breaking the problem down into smaller single-concept pieces can reduce the cognitive load to something manageable.
-
-For example,
-a growing number of educators are including Parsons Problems in their pedagogic repertoire~\citep{parsons,morrison-parsons}.
-Rather than writing programs from scratch,
-learners are given the lines of code they need to solve a problem,
-but in jumbled order.
-Re-ordering them to solve the problem correctly allows them to concentrate on mastering control flow
-without having to devote mental energy to recalling syntax
-or the specifics of library functions.
+but some learners may associate them with violence and racial or gender stereotypes, or simply find them unenjoyable.
 
@@
 
-This principle sounds tautological, but it is easily forgotten.
+This principle is tautological, but it is easily forgotten.
 Novices program differently than experts,
@@
 
+\section{Don't just code}\label{not-just-code}
+
+Our final rule for teaching programming is that you don't have to program to do it.
+Between syntax, semantics, algorithms, and design,
+examples that seem small to instructors can easily overwhelm novices.
+Breaking the problem down into smaller single-concept pieces can reduce the cognitive load to something manageable.
+
+For example,
+a growing number of educators are including Parsons Problems in their pedagogic repertoire~\citep{parsons,morrison-parsons}.
+Rather than writing programs from scratch,
+learners are given the lines of code they need to solve a problem,
+but in jumbled order.
+Re-ordering them to solve the problem correctly allows them to concentrate on mastering control flow
+without having to devote mental energy to recalling syntax
+or the specifics of library functions.
+
 \section*{Conclusion}
@@
 \bibliographystyle{abbrvnat}
-\bibliography{refs}
+\bibliography{paper}
 

### 2017-07-07 paper.tex
Updating README

@@
 \citet{guzdial-myths} refers to the belief that some people are born programmers and others aren't
-as ``computing's most enduring and damaging myth.''  This is often ``confirmed'' by looking at university grade distributions, which are commonly held to be bimodal: a low-scoring hump of those that will never get it, and a high-scoring hump of those that have the right stuff.
+as ``computing's most enduring and damaging myth.''
+This is often ``confirmed'' by looking at university grade distributions,
+which are commonly held to be bimodal:
+a low-scoring hump of those that will never get it
+and a high-scoring hump of those that have the right stuff.
 Our first and most important rule is that this is wrong:
@@
 
-One-on-one tutoring is perhaps the ideal form of teaching: all of a teacher's attention can be focused on one student, and they can completely customise their teaching for their single student, and tailor individual feedback and corrections based on two-way dialogue with that student.
-In realistic settings however, one teacher must teach many students at once, whether that is a handful, tens or even hundreds in large lectures.
-In larger settings, how can teachers possibly hope to clear up learners' many different misconceptions in a reasonable time?
+One-on-one tutoring is the ideal form of teaching:
+all of a teacher's attention can be focused on one student,
+and they can completely customise their teaching for that person
+and tailor individual feedback and corrections based on two-way dialogue with them.
+In realistic settings,
+however,
+one teacher must teach a handful, tens, or hundreds of students at once.
+How can teachers possibly hope to clear up many learners' different misconceptions
+in these larger settings in a reasonable time?
 
@@
 \item The instructor then gives learners a multiple choice question
-   that probes for misconceptions rather than simple factual recall.  A programming example is given in figure \ref{fig:mcq} which relates to integer comparison and loops.
+  that probes for misconceptions rather than simple factual recall.
+  A programming example is given in figure \ref{fig:mcq}
+  which relates to integer comparison and loops.
 
@@
 \begin{verbatim}
-for (int i = 1; i < 10; i++)
-{
-     if (i < 3 || i > 8)
-     {
+for (int i = 1; i < 10; i++) {
+    if (i < 3 || i > 8) {
         System.out.println("Yes");
@@
 \label{fig:mcq}
-\caption{An example multiple choice question, probing learners' understanding of loops and integer comparisons.}
+\caption{An example multiple choice question probing learners' understanding of loops and integer comparisons.}
 \end{figure}
@@
 
-\item It facilitates unintended knowledge transfer: students learn more
-  than the instructor consciously intends to teach by watching \textit{how}
-  instructors do things.  The extra knowledge may be high-level (e.g. whether a program is written top-down or bottom-up) or fairly low-level (e.g. using keyboard shortcuts).
+\item It facilitates unintended knowledge transfer: students learn
+  more than the instructor consciously intends to teach by watching
+  \textit{how} instructors do things.  The extra knowledge may be
+  high-level (e.g., whether a program is written top-down or
+  bottom-up) or fairly low-level (e.g., using keyboard shortcuts).
 
-\item It slows the instructor down: if the instructor has to type in the program
-  as they go along, they can only go twice as fast as their
-  learners, rather than ten-fold faster as they could with slides.
+\item It slows the instructor down: if the instructor has to type in
+  the program as they go along, they can only go twice as fast as
+  their learners, rather than ten-fold faster as they could with
+  slides.
 
@@
 
-\item Watching instructors make mistakes shows learners that it's alright to make mistakes of their own~\citep{barker-live-coding}.
+\item Watching instructors make mistakes shows learners that it's
+  alright to make mistakes of their own~\citep{barker-live-coding}.
   Most people model the behavior of their teachers: if the instructor
@@
 
-\item Instructors can spend too much time typing in boilerplate code that is needed by the lesson, but not
-  directly relevant to it (such as library import statements), can
-  distract learners from the intended thrust of a lesson.  As
-  \citet{willingham-book} says, ``Memory is the residue of thought'';
-  if the instructor spends their time typing boilerplate, that may be
-  all that learners take away.  This can be avoided by starting with a partial skeleton that includes the boilerplate, or having it on hand to copy and paste at the appropriate juncture.
+\item Instructors can spend too much time typing in boilerplate code
+  that is needed by the lesson, but not directly relevant to it (such
+  as library import statements), can distract learners from the
+  intended thrust of a lesson.  As \citet{willingham-book} says,
+  ``Memory is the residue of thought''; if the instructor spends their
+  time typing boilerplate, that may be all that learners take away.
+  This can be avoided by starting with a partial skeleton that
+  includes the boilerplate, or having it on hand to copy and paste
+  when needed.  (Of the two, we prefer the former, since learners
+  may not be able to keep up with copying and pasting.)
 
@@
 
-When instructors are using live coding,
-they usually run the program several times during its development to show what it does.
-Research from peer instruction in physics education shows that
-learners who observe a demonstration \textit{do not} learn better than
-those who did not see the demonstration~\citep{crouch-demo},
-and in fact many learners misremember the outcome of demonstrations afterwards~\citep{miller-demo}.  That is, demonstrations can be useless or actively harmful to learning.
+When instructors are using live coding, they usually run the program
+several times during its development to show what it does.  Research
+from peer instruction in physics education shows that learners who
+observe a demonstration \textit{do not} learn better than those who
+did not see the demonstration~\citep{crouch-demo}, and in fact many
+learners misremember the outcome of demonstrations
+afterwards~\citep{miller-demo}.  In other words, demonstrations can
+actually be useless or actively harmful.
 
@@
 finding the first surname alphabetically from a list of names
-may seem like quite different problems to them, even though more advanced programmers would recognise them as isomorphic.
+may seem like quite different problems to them,
+even though more advanced programmers would recognise them as isomorphic.
 
@@
 
-A principle that applies across all areas of education is that transference only comes with mastery \citep{gick-holyoak}.
+A principle that applies across all areas of education is that
+transference only comes with mastery \citep{gick-holyoak}.
 Courses should therefore stick to one language until learners have progressed far enough with it
@@
 many educators use computer games as a motivating example for programming classes,
-but some learners may associate them with violence and racial or gender stereotypes, or simply find them unenjoyable.
+but some learners may associate them with violence and racial or gender stereotypes,
+or simply find them unenjoyable.
 
@@
 examples that seem small to instructors can easily overwhelm novices.
-Breaking the problem down into smaller single-concept pieces can reduce the cognitive load to something manageable.
+Breaking the problem down into smaller single-concept pieces
+can reduce the cognitive load to something manageable.
 
 For example,
-a growing number of educators are including Parsons Problems in their pedagogic repertoire~\citep{parsons,morrison-parsons}.
+a growing number of educators are including Parsons Problems
+in their pedagogic repertoire~\citep{parsons,morrison-parsons}.
 Rather than writing programs from scratch,
@@
 but in jumbled order.
-Re-ordering them to solve the problem correctly allows them to concentrate on mastering control flow
+Re-ordering them to solve the problem correctly
+allows them to concentrate on mastering control flow
 without having to devote mental energy to recalling syntax

### 2017-07-09 paper.tex
Adding figure and reference

@@
 formal or informal.
-
-% FIXME: mention Repenning's boring/exciting X hard/easy framework from "Moving Beyond Syntax" paper.
 
@@
 
-\begin{figure}\sffamily
+\begin{figure}
 \begin{verbatim}
@@
 
-\citet{morrison-subgoals,morrison-parsons} have shown that students perform better
+\citet{morrison-subgoals,morrison-parsons,margulieux} have shown that students perform better
 when worked examples are broken down into steps (or subgoals)
-which are given names (or labels).
+which are given names (or labels) (Figure~\ref{fig:subgoal}).
 Subgoal labels provide a structure which students can apply to future tasks that they attempt themselves,
 and allow them to communicate with their peers and instructors more efficiently.
-TODO give an example in a figure.
 
-% FIXME give a short code snippet here with some subgoal labels to illustrate - yes please
+\begin{figure}
+
+\noindent
+\textbf{Conventional Materials}
+\begin{verbatim}
+1. Click on "My Blocks" to see the blocks for components you created.
+2. Click on "clap"
+3. Drag out a when clap.Touched block
+4. Click on “clapSound”
+5. Drag out call clapSound.Play
+6. Connect it after when clap.Touched
+\end{verbatim}
+
+\noindent
+\textbf{Subgoal Labeled Materials}
+\begin{verbatim}
+Handle Events from My Blocks
+
+1. Click on "My Blocks" to see the blocks for components you created.
+2. Click on "clap"
+3. Drag out a when clap.Touched block
+
+Set Output from My Blocks
+
+4. Click on “clapSound” and
+5. Drag out call clapSound.Play
+6. Connect it after when clap.Touched
+\end{verbatim}
+\label{fig:subgoal}
+\caption{An example of subgoal labeling (taken from \cite{margulieux})}
+\end{figure}
 

### 2017-07-09 paper.tex
Updating bibliography

@@
 it has been studied extensively in a wide variety of contexts,
-including programming \citep{cutts,porter}.
+including programming \citep{porter-multi,porter-success}.
 In simplified form,
@@
 
-A classic question in computing (and mathematics) education is whether problems are better 
-with context (e.g. find the highest student grade) or without (e.g. find the maximum of the list of numbers).
-\citet{bouvier-context} studied this with a multi-university study and found no difference between
-the two.  They suggest that since it makes no difference, other considerations (such as motivation)
-should be primary.
+A classic question in computing (and mathematics) education
+is whether problems are better with context (e.g., find the highest student grade)
+or without (e.g. find the maximum of the list of numbers).
+\citet{bouvier-context} examined this with a multi-university study
+and found no difference between the two.
+They suggest that since it makes no difference,
+other considerations (such as motivation) should be primary.
 
@@
 or simply find them unenjoyable.
+Whatever examples are chosen,
+the goal must be to move learners as quickly as possible
+from ``hard and boring'' to ``easy and exciting'' \cite{repenning}.
 
-\section{Novices are not experts}\label{not-experts}
+\section{Remember that novices are not experts}\label{not-experts}
 

### 2017-07-10 paper.tex
One more pass through the paper.

@@
 \usepackage{hyperref}
+\usepackage{mdframed}
 
 \title{Ten Simple Rules for Teaching Programming}
-\author{Neil C. C. Brown\footnote{King's College London / neil.c.c.brown@kcl.ac.uk} and Greg Wilson\footnote{Rangle.io / gvwilson@third-bit.com}}
+\author{Greg Wilson\footnote{Rangle.io / gvwilson@third-bit.com} and Neil C. C. Brown\footnote{King's College London / neil.c.c.brown@kcl.ac.uk}}
 
@@
 
-One-on-one tutoring is the ideal form of teaching:
+One-on-one tutoring is perhaps the ideal form of teaching:
 all of a teacher's attention can be focused on one student,
@@
 however,
-one teacher must teach a handful, tens, or hundreds of students at once.
+one teacher must usually teach several, tens, or even hundreds of students at once.
 How can teachers possibly hope to clear up many learners' different misconceptions
@@
   A programming example is given in figure \ref{fig:mcq}
-  which relates to integer comparison and loops.
+  which relates to integer comparison and loops.  There is no point asking a trivial question
+  that all students will get right.  The ideal questions are those where students are most likely to vote equally for each of the options.
 
@@
 \begin{figure}
+\begin{mdframed}
 \begin{verbatim}
 for (int i = 1; i < 10; i++) {
-    if (i < 3 || i > 8) {
+    if (i < 3 || i >= 8) {
         System.out.println("Yes");
@@
 a) 10\\
-b) 9\\
+b) 5\\
 c) 4\\
@@
 \caption{An example multiple choice question probing learners' understanding of loops and integer comparisons.}
+\end{mdframed}
 \end{figure}
@@
 
-A good idea when teaching is
-to provide many similar examples for learners to practice on.
+Instructors usually provide many similar programming examples for learners to practice on.
 But since learners are novices,
@@
 when worked examples are broken down into steps (or subgoals)
-which are given names (or labels) (Figure~\ref{fig:subgoal}).
+which are given names (or labels) -- an example is given in figure~\ref{fig:subgoal}.
 Subgoal labels provide a structure which students can apply to future tasks that they attempt themselves,
@@
 \begin{figure}
-
+\begin{mdframed}
 \noindent
@@
 \caption{An example of subgoal labeling (taken from \cite{margulieux})}
+\end{mdframed}
 \end{figure}
@@
 the goal must be to move learners as quickly as possible
-from ``hard and boring'' to ``easy and exciting'' \cite{repenning}.
+from ``hard and boring'' to ``easy and exciting'' \citep{repenning}.
 
@@
 we must take care to use small, self-contained tasks at a level suitable for novices,
-with tools that suit their needs.
+with tools that suit their needs, and without scoffing.
 
@@
 and are learning more each year.
-Venues like SIGCSE (\url{http://sigcse.org/}) and ITiCSE (\url{http://iticse.acm.org/})
+Venues like SIGCSE (\url{http://sigcse.org/}), ITiCSE (\url{http://iticse.acm.org/}) and ICER (\url{https://icer.hosting.acm.org})
 present a growing number of rigorous, insightful studies
@@
 
+We offer one final observation: do not forget the human element.
+Programmers have a reputation for pouring scorn on certain programming languages (e.g.
+PHP), or for gatekeeping (e.g.
+stating that you can't learn programming if you didn't start young).
+If you are teaching someone to program, the last thing you want to do is make them feel like they can't succeed or that any existing skill (perhaps in PHP) they have is worthless.
+Make your learners feel that they can be a programmer, and they just might become one.
+
 \bibliographystyle{abbrvnat}

### 2017-07-10 paper.tex
Copying latest changes to README

@@
 \title{Ten Simple Rules for Teaching Programming}
-\author{Greg Wilson\footnote{Rangle.io / gvwilson@third-bit.com} and Neil C. C. Brown\footnote{King's College London / neil.c.c.brown@kcl.ac.uk}}
+\author{Neil C. C. Brown\footnote{King's College London / neil.c.c.brown@kcl.ac.uk} and Greg Wilson\footnote{Rangle.io / gvwilson@third-bit.com}}
 
@@
   A programming example is given in figure \ref{fig:mcq}
-  which relates to integer comparison and loops.  There is no point asking a trivial question
-  that all students will get right.  The ideal questions are those where students are most likely to vote equally for each of the options.
+  which relates to integer comparison and loops.
+  There is no point asking a trivial question that all students will get right.
+  The ideal questions are those where students are most likely to vote equally for each of the options.
 
@@
 We offer one final observation: do not forget the human element.
-Programmers have a reputation for pouring scorn on certain programming languages (e.g.
-PHP), or for gatekeeping (e.g.
-stating that you can't learn programming if you didn't start young).
-If you are teaching someone to program, the last thing you want to do is make them feel like they can't succeed or that any existing skill (perhaps in PHP) they have is worthless.
-Make your learners feel that they can be a programmer, and they just might become one.
+Programmers have a reputation for pouring scorn on certain programming languages (e.g., PHP),
+or for gatekeeping (e.g., stating that you can't learn programming if you didn't start young).
+If you are teaching someone to program,
+the last thing you want to do is make them feel like they can't succeed
+or that any existing skill they have (no matter when or how acquired) is worthless.
+Make your learners feel that they can be a programmer,
+and they just might become one.
 

### 2017-07-11 paper.tex
Formatting with PLOS style

@@
-\documentclass{article}
-\usepackage[a4paper]{geometry}
-\usepackage{natbib}
-\usepackage{hyperref}
+% Template for PLoS
+% Version 3.4 January 2017
+%
+% % % % % % % % % % % % % % % % % % % % % %
+%
+% -- IMPORTANT NOTE
+%
+% This template contains comments intended
+% to minimize problems and delays during our production
+% process. Please follow the template instructions
+% whenever possible.
+%
+% % % % % % % % % % % % % % % % % % % % % % %
+%
+% Once your paper is accepted for publication,
+% PLEASE REMOVE ALL TRACKED CHANGES in this file
+% and leave only the final text of your manuscript.
+% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
+% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
+%
+%
+% There are no restrictions on package use within the LaTeX files except that
+% no packages listed in the template may be deleted.
+%
+% Please do not include colors or graphics in the text.
+%
+% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
+%
+% % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- FIGURES AND TABLES
+%
+% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
+%
+% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
+% - Figures should be uploaded separately from your manuscript file.
+% - Figures generated using LaTeX should be extracted and removed from the PDF before submission.
+% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
+% For figure citations, please use "Fig" instead of "Figure".
+% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
+%
+% Tables should be cell-based and may not contain:
+% - spacing/line breaks within cells to alter layout or alignment
+% - do not nest tabular environments (no tabular environments within tabular environments)
+% - no graphics or colored text (cell background color/shading OK)
+% See http://journals.plos.org/plosone/s/tables for table guidelines.
+%
+% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
+%
+% IMPORTANT
+% Below are a few tips to help format your equations and other special characters according to our specifications.
+For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
+%
+% For inline equations, please be sure to include all portions of an equation in the math environment.
+For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
+%
+% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
+%
+% Please add line breaks to long display equations when possible in order to fit size of the column.
+%
+% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
+%
+% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2".
+%
+% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+%
+% Please contact latex@plos.org with any questions.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+
+\documentclass[10pt,letterpaper]{article}
+\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
+
+% amsmath and amssymb packages, useful for mathematical formulas and symbols
+\usepackage{amsmath,amssymb}
+
+% Use adjustwidth environment to exceed column width (see example table in text)
+\usepackage{changepage}
+
+% Use Unicode characters when possible
+\usepackage[utf8x]{inputenc}
+
+% textcomp package and marvosym package for additional characters
+\usepackage{textcomp,marvosym}
+
+% cite package, to clean up citations in the main text. Do not remove.
+\usepackage{cite}
+
+% Use nameref to cite supporting information files (see Supporting Information section for more info)
+% \usepackage{nameref,hyperref}
+
+% line numbers
+\usepackage[right]{lineno}
+
+% ligatures disabled
+\usepackage{microtype}
+\DisableLigatures[f]{encoding = *, family = * }
+
+% color can be used to apply background shading to table cells only
+\usepackage[table]{xcolor}
+
+% array package and thick rules for tables
+\usepackage{array}
+
+% create "+" rule type for thick vertical lines
+\newcolumntype{+}{!{\vrule width 2pt}}
+
+% frames around boxes
 \usepackage{mdframed}
 
-\title{Ten Simple Rules for Teaching Programming}
-\author{Neil C. C. Brown\footnote{King's College London / neil.c.c.brown@kcl.ac.uk} and Greg Wilson\footnote{Rangle.io / gvwilson@third-bit.com}}
+% create \thickcline for thick horizontal lines of variable length
+\newlength\savedwidth
+\newcommand\thickcline[1]{%
+  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+  \cline{#1}%
+  \noalign{\vskip\arrayrulewidth}%
+  \noalign{\global\arrayrulewidth\savedwidth}%
+}
+
+% \thickhline command for thick horizontal lines that span the table
+\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+\hline
+\noalign{\global\arrayrulewidth\savedwidth}}
+
+% Remove comment for double spacing
+%\usepackage{setspace}
+%\doublespacing
+
+% Text layout
+\raggedright
+\setlength{\parindent}{0.5cm}
+\textwidth 5.25in
+\textheight 8.75in
+
+% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
+% Captions will be left justified
+\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
+\renewcommand{\figurename}{Fig}
+
+% Use the PLoS provided BiBTeX style
+\bibliographystyle{plos2015}
+
+% Remove brackets from numbering in List of References
+\makeatletter
+\renewcommand{\@biblabel}[1]{\quad#1.}
+\makeatother
+
+% Leave date blank
+\date{}
+
+% Header and Footer with logo
+\usepackage{lastpage,fancyhdr,graphicx}
+\usepackage{epstopdf}
+\pagestyle{myheadings}
+\pagestyle{fancy}
+\fancyhf{}
+\setlength{\headheight}{27.023pt}
+% \lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
+\rfoot{\thepage/\pageref{LastPage}}
+\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
+\fancyheadoffset[L]{2.25in}
+\fancyfootoffset[L]{2.25in}
+% \lfoot{\sf PLOS}
+
+%% Include all macros below
+
+\newcommand{\rulemajor}[2]{\section{#1}\label{#2}}
+\newcommand{\ruleref}[1]{Rule~\ref{#1}}
+\newcommand{\url}[1]{\texttt{#1}}
+
+%% END MACROS SECTION
 
 \begin{document}
+\vspace*{0.2in}
 
-\maketitle
+% Title must be 250 characters or less.
+\begin{flushleft}
+{\Large
+\textbf\newline{Ten simple rules for teaching programming} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+}
+\newline
+% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
+\\
+{Neil C.C.\ Brown}\textsuperscript{1{\ddag}*},
+{Greg~Wilson}\textsuperscript{2{\ddag}*}
+\\
+\bigskip
+\textbf{1} King's College London / neil.c.c.brown@kcl.ac.uk \\
+\textbf{2} Rangle.io / gvwilson@third-bit.com \\
+
+\bigskip
+
+% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
+%
+% Remove or comment out the author notes below if they aren't used.
+%
+% Primary Equal Contribution Note
+{\ddag} These authors contributed equally to this work.
+
+% Additional Equal Contribution Note
+% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
+% \ddag These authors also contributed equally to this work.
+
+% Current address notes
+% \textcurrency Current Address: Dept/Program/Center, Institution Name, City, State, Country % change symbol to "\textcurrency a" if more than one current address note
+% \textcurrency b Insert second current address
+% \textcurrency c Insert third current address
+
+% Deceased author note
+% \dag Deceased
+
+% Group/Consortium Author Note
+% \textpilcrow Membership list can be found in the Acknowledgments section.
+
+% Use the asterisk to denote corresponding authorship and provide email address in note below.
+* corresponding author
+
+\end{flushleft}
+
+% Please keep the abstract below 300 words
+
+% \linenumbers
+
+% \section*{Abstract}
+% FIXME write abstrat
+
+% Please keep the Author Summary between 150 and 200 words
+% Use first person. PLOS ONE authors please skip this step.
+% Author Summary not valid for PLOS ONE submissions.
+% \section*{Author summary}
+% FIXME write author summary
 
@@
 Research from educational psychology suggests that teaching and learning are subject-specific
-activities~\citep{mayer-subject}: learning programming has a different set of challenges and techniques
+activities~\cite{mayer-subject}: learning programming has a different set of challenges and techniques
 than learning physics or learning to read and write.  Computing is a younger discipline than
@@
 
-\section{There is no geek gene}\label{gene}
+\rulemajor{There is no geek gene}{gene}
 
-\citet{guzdial-myths} refers to the belief that some people are born programmers and others aren't
+\cite{guzdial-myths} refers to the belief that some people are born programmers and others aren't
 as ``computing's most enduring and damaging myth.''
@@
 
-The most powerful evidence for this comes from \citet{patitsas-bimodal}.
+The most powerful evidence for this comes from \cite{patitsas-bimodal}.
 They examined grade distributions in introductory computing courses at a large university,
@@
 Beliefs such as this are known to have powerful effects on education outcomes
-\citep{alvidrez-weinstein,brophy,jussim-eccles}.
+\cite{alvidrez-weinstein,brophy,jussim-eccles}.
 If instructors believe that ``some kids get it and some kids don't'',
@@
 
-\section{Use peer instruction}\label{peer-instruction}
+\rulemajor{Use peer instruction}{peer-instruction}
 
@@
 The best method developed so far for larger-scale classrooms is called Peer Instruction.
-Originally created by Eric Mazur at Harvard~\citep{mazur-peer-instruction},
+Originally created by Eric Mazur at Harvard~\cite{mazur-peer-instruction},
 it has been studied extensively in a wide variety of contexts,
-including programming \citep{porter-multi,porter-success}.
+including programming \cite{porter-multi,porter-success}.
 In simplified form,
@@
   that probes for misconceptions rather than simple factual recall.
-  A programming example is given in figure \ref{fig:mcq}
+  A programming example is given in Figure~\ref{fig:mcq}
   which relates to integer comparison and loops.
@@
 because of varied prior experience).
-Pair programming (Rule~\ref{pair-programming}) can be used to mitigate this.
+Pair programming (\ruleref{pair-programming}) can be used to mitigate this.
 
-\begin{figure}
-\begin{mdframed}
-\begin{verbatim}
-for (int i = 1; i < 10; i++) {
-    if (i < 3 || i >= 8) {
-        System.out.println("Yes");
-    }
-}
-\end{verbatim}
-How many times will the above code print out the word Yes?\\
-a) 10\\
-b) 5\\
-c) 4\\
-d) 3
-\label{fig:mcq}
-\caption{An example multiple choice question probing learners' understanding of loops and integer comparisons.}
-\end{mdframed}
-\end{figure}
-
-\section{Use live coding}\label{live-coding}
+\rulemajor{Use live coding}{live-coding}
 
 Rather than using slides,
-instructors should create programs in front of their learners~\citep{rubin-live-coding}.
+instructors should create programs in front of their learners~\cite{rubin-live-coding}.
 This is more effective for multiple reasons:
@@
 \item Watching instructors make mistakes shows learners that it's
-  alright to make mistakes of their own~\citep{barker-live-coding}.
+  alright to make mistakes of their own~\cite{barker-live-coding}.
   Most people model the behavior of their teachers: if the instructor
@@
   as library import statements), can distract learners from the
-  intended thrust of a lesson.  As \citet{willingham-book} says,
+  intended thrust of a lesson.  As \cite{willingham-book} says,
   ``Memory is the residue of thought''; if the instructor spends their
@@
 
-\section{Have students make predictions}\label{predictions}
+\rulemajor{Have students make predictions}{predictions}
 
@@
 observe a demonstration \textit{do not} learn better than those who
-did not see the demonstration~\citep{crouch-demo}, and in fact many
+did not see the demonstration~\cite{crouch-demo}, and in fact many
 learners misremember the outcome of demonstrations
-afterwards~\citep{miller-demo}.  In other words, demonstrations can
+afterwards~\cite{miller-demo}.  In other words, demonstrations can
 actually be useless or actively harmful.
@@
 
-\section{Use pair programming}\label{pair-programming}
+\rulemajor{Use pair programming}{pair-programming}
 
@@
 The two switch roles several times per hour.
-Pair programming is a good practice in real-life programming \citep{hannay},
-and also a good way to teach \citep{mcdowell}.
+Pair programming is a good practice in real-life programming \cite{hannay},
+and also a good way to teach \cite{mcdowell}.
 Partners can not only help each other out during practical exercises,
@@
 
-\section{Use worked examples with labelled subgoals}\label{worked-examples}
+\rulemajor{Use worked examples with labelled subgoals}{worked-examples}
 
@@
 
-\citet{morrison-subgoals,morrison-parsons,margulieux} have shown that students perform better
+\cite{morrison-subgoals,morrison-parsons,margulieux} have shown that students perform better
 when worked examples are broken down into steps (or subgoals)
-which are given names (or labels) -- an example is given in figure~\ref{fig:subgoal}.
+which are given names (or labels) -- an example is given in Figure~\ref{fig:subgoal}.
 Subgoal labels provide a structure which students can apply to future tasks that they attempt themselves,
@@
 
-\begin{figure}
-\begin{mdframed}
-\noindent
-\textbf{Conventional Materials}
-\begin{verbatim}
-1. Click on "My Blocks" to see the blocks for components you created.
-2. Click on "clap"
-3. Drag out a when clap.Touched block
-4. Click on “clapSound”
-5. Drag out call clapSound.Play
-6. Connect it after when clap.Touched
-\end{verbatim}
-
-\noindent
-\textbf{Subgoal Labeled Materials}
-\begin{verbatim}
-Handle Events from My Blocks
-
-1. Click on "My Blocks" to see the blocks for components you created.
-2. Click on "clap"
-3. Drag out a when clap.Touched block
-
-Set Output from My Blocks
-
-4. Click on “clapSound” and
-5. Drag out call clapSound.Play
-6. Connect it after when clap.Touched
-\end{verbatim}
-\label{fig:subgoal}
-\caption{An example of subgoal labeling (taken from \cite{margulieux})}
-\end{mdframed}
-\end{figure}
-
-\section{Stick to one language}\label{one-language}
+\rulemajor{Stick to one language}{one-language}
 
 A principle that applies across all areas of education is that
-transference only comes with mastery \citep{gick-holyoak}.
+transference only comes with mastery \cite{gick-holyoak}.
 Courses should therefore stick to one language until learners have progressed far enough with it
@@
 
-\section{Use authentic tasks}\label{authentic-tasks}
+\rulemajor{Use authentic tasks}{authentic-tasks}
 
@@
 learners remembered more of the material when re-tested after a delay,
-and were more likely to stay in computing programs \citep{guzdial-media}.
+and were more likely to stay in computing programs \cite{guzdial-media}.
 This is a particular instance of a larger observation:
@@
 or without (e.g. find the maximum of the list of numbers).
-\citet{bouvier-context} examined this with a multi-university study
+\cite{bouvier-context} examined this with a multi-university study
 and found no difference between the two.
@@
 the goal must be to move learners as quickly as possible
-from ``hard and boring'' to ``easy and exciting'' \citep{repenning}.
+from ``hard and boring'' to ``easy and exciting'' \cite{repenning}.
 
-\section{Remember that novices are not experts}\label{not-experts}
+\rulemajor{Remember that novices are not experts}{not-experts}
 
@@
 
-\section{Don't just code}\label{not-just-code}
+\rulemajor{Don't just code}{not-just-code}
 
@@
 a growing number of educators are including Parsons Problems
-in their pedagogic repertoire~\citep{parsons,morrison-parsons}.
+in their pedagogic repertoire~\cite{parsons,morrison-parsons}.
 Rather than writing programs from scratch,
@@
 
-\bibliographystyle{abbrvnat}
+\pagebreak
+
+\begin{figure}[h]
+\begin{mdframed}
+\begin{verbatim}
+for (int i = 1; i < 10; i++) {
+    if (i < 3 || i >= 8) {
+        System.out.println("Yes");
+    }
+}
+\end{verbatim}
+How many times will the above code print out the word Yes?\\
+a) 10\\
+b) 5\\
+c) 4\\
+d) 3
+\end{mdframed}
+\caption{An example multiple choice question probing learners' understanding of loops and integer comparisons.}
+\label{fig:mcq}
+\end{figure}
+
+\begin{figure}[h]
+\begin{mdframed}
+\noindent
+\textbf{Conventional Materials}
+\begin{verbatim}
+1. Click on "My Blocks" to see the blocks for components you created.
+2. Click on "clap"
+3. Drag out a when clap.Touched block
+4. Click on “clapSound”
+5. Drag out call clapSound.Play
+6. Connect it after when clap.Touched
+\end{verbatim}
+
+\noindent
+\textbf{Subgoal Labeled Materials}
+\begin{verbatim}
+Handle Events from My Blocks
+
+1. Click on "My Blocks" to see the blocks for components you created.
+2. Click on "clap"
+3. Drag out a when clap.Touched block
+
+Set Output from My Blocks
+
+4. Click on "clapSound" and
+5. Drag out call clapSound.Play
+6. Connect it after when clap.Touched
+\end{verbatim}
+\end{mdframed}
+\caption{An example of subgoal labeling (taken from \cite{margulieux})}
+\label{fig:subgoal}
+\end{figure}
+
+% \nolinenumbers
+
+% Either type in your references using
+% \begin{thebibliography}{}
+% \bibitem{}
+% Text
+% \end{thebibliography}
+%
+% or
+%
+% Compile your BiBTeX database using our plos2015.bst
+% style file and paste the contents of your .bbl file
+% here. See http://journals.plos.org/plosone/s/latex for
+% step-by-step instructions.
+
+\pagebreak
+
 \bibliography{paper}

### 2017-07-21 paper.tex
Fixed up places in the text where we used to use \citet, but now must use author name~\cite

@@
 
-\cite{guzdial-myths} refers to the belief that some people are born programmers and others aren't
+Guzdial~\cite{guzdial-myths} refers to the belief that some people are born programmers and others aren't
 as ``computing's most enduring and damaging myth.''
@@
 
-The most powerful evidence for this comes from \cite{patitsas-bimodal}.
+The most powerful evidence for this comes from Patitsas et al.~\cite{patitsas-bimodal}.
 They examined grade distributions in introductory computing courses at a large university,
@@
   as library import statements), can distract learners from the
-  intended thrust of a lesson.  As \cite{willingham-book} says,
+  intended thrust of a lesson.  As Willingham~\cite{willingham-book} says,
   ``Memory is the residue of thought''; if the instructor spends their
@@
 
-\cite{morrison-subgoals,morrison-parsons,margulieux} have shown that students perform better
+Morrison et al.~\cite{morrison-subgoals,morrison-parsons,margulieux} have shown that students perform better
 when worked examples are broken down into steps (or subgoals)
@@
 learners remembered more of the material when re-tested after a delay,
-and were more likely to stay in computing programs \cite{guzdial-media}.
+and were more likely to stay in computing programs~\cite{guzdial-media}.
 This is a particular instance of a larger observation:
@@
 or without (e.g. find the maximum of the list of numbers).
-\cite{bouvier-context} examined this with a multi-university study
+Bouvier et al.~\cite{bouvier-context} examined this with a multi-university study
 and found no difference between the two.

### 2017-07-21 paper.tex
Adjusted according to Barbara's comments.

@@
   which relates to integer comparison and loops.
-  There is no point asking a trivial question that all students will get right.
-  The ideal questions are those where students are most likely to vote equally for each of the options.
+  There is no point asking a trivial question that all students will get right or one with wrong answers which no student will pick.
+  The ideal questions are those where 40--60\% of students are likely to get the right answer first time, and where every wrong answer corresponds to a misconception and will be picked by at least some students~\cite[p23]{nrc-reaching}.
 
@@
 
+It is also useful for motivation to make sure that students can accomplish a visible result
+quickly.  Providing pre-written software libraries or source code can help to start students
+closer to a meaningful result.  The idea that students must write all the code they need
+themselves is the relic of a bygone era of home microcomputers (and it was not true even then).
+Pick the task that you actually want to the students to engage in, and provide everything else.
+
 A classic question in computing (and mathematics) education
@@
 without having to devote mental energy to recalling syntax
-or the specifics of library functions.
+or the specifics of library functions.  They are also liked by learners; Ericson et al.~\cite{ericson-ebook} found
+that learners were more likely to attempt Parsons Problems than nearby multiple choice questions.
 
@@
 and they just might become one.
+
+\section*{Acknowledgments}
+
+We are grateful to Barbara Ericson for her comments on a draft of this article.
 

### 2017-07-21 paper.tex
Added a reference to rule 9.
One more pass through the paper; various minor changes.

--- 2017-07-21 paper.tex
@@
 mathematics, physics, or biology, and while there have been correspondingly fewer studies of how best to teach it,
-there is a growing body of evidence about what works and doesn't.
+there is a growing body of evidence about what works and what doesn't.
 This paper presents ten simple rules that should be the foundation of any teaching of programming, whether
@@
 and that those faculty were even more likely to see the distributions as bimodal
-if they believed some students are innately predisposed to do well in computer science.
+if they believed that some students are innately predisposed to do well in computer science.
 
@@
 they will (consciously or unconsciously) invest less in those whom they put in the second category.
-When combined with cultural stereotypes about who is and isn't a "natural programmer",
+When combined with cultural stereotypes about who is and isn't a ``natural programmer'',
 the downward spiral of under-achievement that results from differential attention
@@
   A programming example is given in Figure~\ref{fig:mcq}
-  which relates to integer comparison and loops.
-  There is no point asking a trivial question that all students will get right.
-  The ideal questions are those where students are most likely to vote equally for each of the options.
+  which relates to integer comparison and loops.  The multiple choice question must be well designed.
+  There is no point asking a trivial question that all students will get right or one with meaningless wrong answers which no student will pick.
+  The ideal questions are those where 40--60\% of students are likely to get the right answer first time~\cite[p23]{nrc-reaching}, and where every wrong answer corresponds to a misconception that will cause it to be picked by at least some students.
 
@@
   \begin{itemize}
-  \item If they all have the right answer, move on.
+  \item If they all have the right answer, the instructor can move on.
   \item If they all have the same wrong answer,
@@
   allows instructors to go off road and follow their learners'
-  interests.
+  interests or answer unanticipated questions.
 
@@
   high-level (e.g., whether a program is written top-down or
-  bottom-up) or fairly low-level (e.g., using keyboard shortcuts).
+  bottom-up) or fairly low-level (e.g., learning useful editor shortcuts).
 
@@
   their learners, rather than ten-fold faster as they could with
-  slides.
+  slides -- which risks leaving everyone behind.
 
@@
   that is needed by the lesson, but not directly relevant to it (such
-  as library import statements), can distract learners from the
+  as library import statements). Not only does this slow things down, it can distract learners from the
   intended thrust of a lesson.  As Willingham~\cite{willingham-book} says,
@@
 When instructors are using live coding, they usually run the program
-several times during its development to show what it does.  Research
+several times during its development to show what it does.  Surprising research
 from peer instruction in physics education shows that learners who
-observe a demonstration \textit{do not} learn better than those who
+observe a demonstration \textit{do not learn better} than those who
 did not see the demonstration~\cite{crouch-demo}, and in fact many
@@
 by holding up a cue card (A/B/C/D),
-or by talking to their neighbour.
+or by talking to their neighbor.
 We speculate that the sting of being publicly wrong leads learners to pay more attention
@@
 instructors should be careful not to punish or criticise students who predicted wrongly,
-but rather to use those incorrect predictions as a spur to further exploration.
+but rather to use those incorrect predictions as a spur to further exploration and explanation.
 
@@
 finding the first surname alphabetically from a list of names
-may seem like quite different problems to them,
+may seem like quite different problems to learners,
 even though more advanced programmers would recognise them as isomorphic.
@@
 which are given names (or labels) -- an example is given in Figure~\ref{fig:subgoal}.
-Subgoal labels provide a structure which students can apply to future tasks that they attempt themselves,
-and allow them to communicate with their peers and instructors more efficiently.
+Subgoal labels provide a structure which allow learners to see the similarities between coding problems and to communicate with their peers and instructors more efficiently.
+Learners can then apply the labels to future tasks that they attempt themselves.
 
@@
 
-One caution is that context can inadvertently exclude some people while drawing others in.
+One caution about choosing context is that context can inadvertently exclude some people while drawing others in.
 For example,
@@
 
+To help students accomplish a visible and satisfting result
+quickly, instructors can provide some pre-written software libraries or source code that start students
+closer to the end goal.  The idea that students must start from scratch and write all the code they need
+themselves is the relic of a bygone era of home microcomputers (and it was not true even then).
+Pick the task that you actually want to the students to engage in, and provide everything else pre-made.
+
 \rulemajor{Remember that novices are not experts}{not-experts}
@@
 This principle is tautological, but it is easily forgotten.
-Novices program differently than experts,
+Novices program differently than experts~\cite{parnin-expertise},
 and need different approaches or tools.
@@
 Our final rule for teaching programming is that you don't have to program to do it.
-Between syntax, semantics, algorithms, and design,
-examples that seem small to instructors can easily overwhelm novices.
+Faced with the challenges of learning syntax, semantics, algorithms, and design,
+examples that seem small to instructors can still easily overwhelm novices.
 Breaking the problem down into smaller single-concept pieces
@@
 without having to devote mental energy to recalling syntax
-or the specifics of library functions.
+or the specifics of library functions.  They are also liked by learners; Ericson et al.~\cite{ericson-ebook} found
+that learners were more likely to attempt Parsons Problems than nearby multiple choice questions in an ebook.
 
@@
 
-Like anything involving human subjects,
+The ten rules presented here are backed up by scientific research.
+Like any research involving human subjects,
 studies of computing education must necessarily be hedged with qualifiers.
@@
 Future work may overturn or qualify some of our ten simple rules,
-but they form a solid basis for any educational effort.
+but they form a solid basis for any educational effort to the best of our current knowledge.
 
@@
 Programmers have a reputation for pouring scorn on certain programming languages (e.g., PHP),
-or for gatekeeping (e.g., stating that you can't learn programming if you didn't start young).
+or for gatekeeping (e.g., stating that you cannot learn programming if you did not start young).
 If you are teaching someone to program,

### 2017-08-10 paper.tex
Replaced simple rules with quick tips

@@
 {\Large
-\textbf\newline{Ten simple rules for teaching programming} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+\textbf\newline{Ten quick tips for teaching programming} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
 }
@@
 there is a growing body of evidence about what works and what doesn't.
-This paper presents ten simple rules that should be the foundation of any teaching of programming, whether
+This paper presents ten quick tips that should be the foundation of any teaching of programming, whether
 formal or informal.
@@
 and a high-scoring hump of those that have the right stuff.
-Our first and most important rule is that this is wrong:
+Our first and most important tip is that this is wrong:
 competence at programming is not innate,
@@
 
-Our final rule for teaching programming is that you don't have to program to do it.
+Our final tip for teaching programming is that you don't have to program to do it.
 Faced with the challenges of learning syntax, semantics, algorithms, and design,
@@
 
-The ten rules presented here are backed up by scientific research.
+The ten tips presented here are backed up by scientific research.
 Like any research involving human subjects,
@@
 with immediate practical application.
-Future work may overturn or qualify some of our ten simple rules,
+Future work may overturn or qualify some of our ten tips,
 but they form a solid basis for any educational effort to the best of our current knowledge.

### 2017-08-10 paper.tex
Trialling the idea of adding an evidence strength measure.

@@
 
+% To hide this, toggle the two macros below:
+%\newcommand{\evidence}[1]{}
+\newcommand{\evidence}[1]{\vspace{-0.3cm}\textbf{Evidence strength: #1}\vspace{0.5cm}}
+
 %% END MACROS SECTION
@@
 \rulemajor{There is no geek gene}{gene}
+\evidence{Medium\footnote{Our evidence strength is an informal assessment, factoring in whether the effect has been replicated, and whether it has been shown specifically in computing, or only in other disciplines from which we believe it will transfer.}}
 
@@
 \rulemajor{Use peer instruction}{peer-instruction}
+\evidence{Very high}
 
@@
 \rulemajor{Use live coding}{live-coding}
+\evidence{Low}
 
@@
 \rulemajor{Have students make predictions}{predictions}
+\evidence{Low}
 
@@
 \rulemajor{Use pair programming}{pair-programming}
+\evidence{High}
 
@@
 \rulemajor{Use worked examples with labelled subgoals}{worked-examples}
+\evidence{High}
 
@@
 \rulemajor{Stick to one language}{one-language}
+\evidence{Low}
 
@@
 \rulemajor{Use authentic tasks}{authentic-tasks}
+\evidence{Medium}
 
@@
 
-To help students accomplish a visible and satisfting result
+To help students accomplish a visible and satisfying result
 quickly, instructors can provide some pre-written software libraries or source code that start students
@@
 \rulemajor{Remember that novices are not experts}{not-experts}
+\evidence{Medium}
 
@@
 \rulemajor{Don't just code}{not-just-code}
+\evidence{Medium}
 
@@
 
-We are grateful to Barbara Ericson for her comments on a draft of this article.
+We are grateful to Barbara Ericson and Leo Porter for their comments on a draft of this article.
 

### 2017-08-10 paper.tex
Corrected the description of peer instruction to state that the group discussion should always take place.

@@
 In simplified form,
-peer instruction proceeds in three phases:
+peer instruction proceeds in several phases:
 
@@
 
-\item Learners then vote on the answer to the question.
+\item Learners then vote on the answer to the question individually, thus formalising their initial prediction.
 
-  \begin{itemize}
-  \item If they all have the right answer, the instructor can move on.
-  \item If they all have the same wrong answer,
-    the instructor addresses that specific misconception.
-  \item If they have a mix of right and wrong answers,
-    they are given several minutes to discuss those answers with one another
+\item Learners are then given several minutes to discuss those answers with one another
     in small groups (typically 2-4 students)
     and then reconvene and vote again.
+\item Then the instructor can act on the latest answers.
+  \begin{itemize}
+  \item If all the learners have the right answer, the instructor can move on.
+  \item If some of the wrong answers remain popular after group discussion,
+    the instructor can addresses those specific misconceptions directly or engage in class-wide discussion.
   \end{itemize}

### 2017-10-14 paper.tex
Changing affiliation
Addressing reviewers' comments

@@
 % To hide this, toggle the two macros below:
-%\newcommand{\evidence}[1]{}
-\newcommand{\evidence}[1]{\vspace{-0.3cm}\textbf{Evidence strength: #1}\vspace{0.5cm}}
+\newcommand{\evidence}[1]{}
+%\newcommand{\evidence}[1]{\vspace{-0.3cm}\textbf{Evidence strength: #1}\vspace{0.5cm}}
 
@@
 \textbf{1} King's College London / neil.c.c.brown@kcl.ac.uk \\
-\textbf{2} Rangle.io / gvwilson@third-bit.com \\
+\textbf{2} DataCamp / gvwilson@third-bit.com \\
 
@@
 
-\rulemajor{There is no geek gene}{gene}
+These tips will be useful to anyone teaching programming at any level
+and to any audience. A larger list aimed primarily at K-12 audiences
+can be found at \cite{cs-teaching-tips}.
+
+\rulemajor{Remember that there is no geek gene}{gene}
 \evidence{Medium\footnote{Our evidence strength is an informal assessment, factoring in whether the effect has been replicated, and whether it has been shown specifically in computing, or only in other disciplines from which we believe it will transfer.}}
@@
 \rulemajor{Use peer instruction}{peer-instruction}
-\evidence{Very high}
+\evidence{High}
 
@@
 
+Note that live coding does not always have to start with a blank screen:
+instructors may give students some starter code
+that relies solely on concepts they have already mastered,
+and then extend it or modify it with live coding.
+
 \rulemajor{Have students make predictions}{predictions}
@@
 They suggest that since it makes no difference,
-other considerations (such as motivation) should be primary.
+other considerations (such as motivation) should be given priority.
 
@@
 We offer one final observation: do not forget the human element.
-Programmers have a reputation for pouring scorn on certain programming languages (e.g., PHP),
+Programmers have a reputation for pouring scorn on certain programming tools (e.g., pouring scorn on spreadsheets),
 or for gatekeeping (e.g., stating that you cannot learn programming if you did not start young).
@@
 We are grateful to Barbara Ericson and Leo Porter for their comments on a draft of this article.
+
+\pagebreak
+
+\section*{Ten Quick Tips for Teaching Programming}
+
+\begin{enumerate}
+\item Remember that there is no geek gene.
+\item Use peer instruction.
+\item Use live coding.
+\item Have students make predictions.
+\item Use pair programming.
+\item Use worked examples with labelled subgoals.
+\item Stick to one language.
+\item Use authentic tasks.
+\item Remember that novices are not experts.
+\item Don't just code.
+\end{enumerate}
 
@@
 3. Drag out a when clap.Touched block
-4. Click on “clapSound”
+4. Click on "clapSound"
 5. Drag out call clapSound.Play

### 2017-10-31 paper.tex
Swapped out the subgoal labeling figure for a custom figure rather than copying from existing paper.

@@
 
-Morrison et al.~\cite{morrison-subgoals,morrison-parsons,margulieux} have shown that students perform better
+Margulieux and Morrison et al.~\cite{morrison-subgoals,morrison-parsons,margulieux} have shown that students perform better
 when worked examples are broken down into steps (or subgoals)
@@
 
-We are grateful to Barbara Ericson and Leo Porter for their comments on a draft of this article.
+We are grateful to Barbara Ericson and Leo Porter for their comments on a draft of this article, and to Lauren Margulieux for providing the example used in Figure \ref{fig:subgoal}.
 
@@
 \begin{verbatim}
-1. Click on "My Blocks" to see the blocks for components you created.
-2. Click on "clap"
-3. Drag out a when clap.Touched block
-4. Click on "clapSound"
-5. Drag out call clapSound.Play
-6. Connect it after when clap.Touched
+1. Click on "AccelerometerSensor1"
+2. Drag out a when AccelerometerSensor1.AccelerationChanged block
+3. Click on ?cowbellSound? 
+4. Drag out call cowbellSound.Play and connect it after
+      when AccelerometerSensor1.AccelerationChanged
 \end{verbatim}
@@
 Handle Events from My Blocks
-
-1. Click on "My Blocks" to see the blocks for components you created.
-2. Click on "clap"
-3. Drag out a when clap.Touched block
+1. Click on "AccelerometerSensor1"
+2. Drag out a when AccelerometerSensor1.AccelerationChanged block
 
 Set Output from My Blocks
-
-4. Click on "clapSound" and
-5. Drag out call clapSound.Play
-6. Connect it after when clap.Touched
+3. Click on ?cowbellSound? 
+4. Drag out call cowbellSound.Play and connect it after
+      when AccelerometerSensor1.AccelerationChanged
 \end{verbatim}
 \end{mdframed}
-\caption{An example of subgoal labeling (taken from \cite{margulieux})}
+\caption{An example of subgoal labeling.}
 \label{fig:subgoal}

### 2018-02-23 paper.tex
Revising
There are always more changes

@@
 
+% Trying to typeset code
+\usepackage{newfloat}
+\DeclareFloatingEnvironment[
+    fileext=loc,
+    listname={List of Code Samples},
+    name=Code,
+    placement=tbhp,
+    within=none,
+]{code}
+
 % Use the PLoS provided BiBTeX style
@@
 \\
-{Neil C.C.\ Brown}\textsuperscript{1{\ddag}*},
+{Neil C.C.\ Brown}\textsuperscript{1{\ddag}},
 {Greg~Wilson}\textsuperscript{2{\ddag}*}
@@
 \bigskip
-\textbf{1} King's College London / neil.c.c.brown@kcl.ac.uk \\
-\textbf{2} DataCamp / gvwilson@third-bit.com \\
+\textbf{1} Department of Informatics, King's College London, London, UK, neil.c.c.brown@kcl.ac.uk \\
+\textbf{2} DataCamp, Toronto, Onario, Canada, gvwilson@third-bit.com \\
 
@@
   that probes for misconceptions rather than simple factual recall.
-  A programming example is given in Figure~\ref{fig:mcq}
+  A programming example is given in Code~\ref{fig:mcq}
   which relates to integer comparison and loops.  The multiple choice question must be well designed.
@@
 and then extend it or modify it with live coding.
+Instructors who use live coding should ensure that
+learners have reference material available after lectures,
+such as a textbook,
+but should also recognize that students of all ages
+increasingly turn to Q\&A sites such as Stack Overflow for information.
 
@@
 when worked examples are broken down into steps (or subgoals)
-which are given names (or labels) -- an example is given in Figure~\ref{fig:subgoal}.
+which are given names (or labels) -- an example is given in Code~\ref{fig:subgoal}.
 Subgoal labels provide a structure which allow learners to see the similarities between coding problems and to communicate with their peers and instructors more efficiently.
@@
 
-We are grateful to Barbara Ericson and Leo Porter for their comments on a draft of this article, and to Lauren Margulieux for providing the example used in Figure \ref{fig:subgoal}.
+We are grateful to Barbara Ericson and Leo Porter for their comments on a draft of this article,
+and to Lauren Margulieux for providing the example used in Code~\ref{fig:subgoal}.
 
@@
 
-\begin{figure}[h]
+\begin{code}[h]
 \begin{mdframed}
@@
 \label{fig:mcq}
-\end{figure}
+\end{code}
 
-\begin{figure}[h]
+\begin{code}[h]
 \begin{mdframed}
@@
 2. Drag out a when AccelerometerSensor1.AccelerationChanged block
-3. Click on ?cowbellSound? 
+3. Click on "cowbellSound"
 4. Drag out call cowbellSound.Play and connect it after
@@
 Set Output from My Blocks
-3. Click on ?cowbellSound? 
+3. Click on "cowbellSound"
 4. Drag out call cowbellSound.Play and connect it after
@@
 \label{fig:subgoal}
-\end{figure}
+\end{code}
 
@@
 
-\bibliography{paper}
+% \bibliography{paper}
+
+\begin{thebibliography}{10}
+
+\bibitem{mayer-subject}
+Mayer RE.
+\newblock Teaching of Subject Matter.
+\newblock Annual Review of Psychology. 2004;55(1):715--744.
+\newblock doi:{10.1146/annurev.psych.55.082602.133124}.
+
+\bibitem{cs-teaching-tips}
+Lewis C. CS Teaching Tips; 2017.
+\newblock \texttt{http://csteachingtips.org/}.
+
+\bibitem{guzdial-myths}
+Guzdial M. Top 10 Myths About Teaching Computer Science; 2015.
+\newblock
+  \texttt{https://cacm.acm.org/blogs/blog-cacm/189498-top-10-myths-about-teaching-computer-science/fulltext}.
+
+\bibitem{patitsas-bimodal}
+Patitsas E, Berlin J, Craig M, Easterbrook S.
+\newblock Evidence That Computer Science Grades Are Not Bimodal.
+\newblock In: Proceedings of the 2016 ACM Conference on International Computing
+  Education Research. ICER '16. New York, NY, USA: ACM; 2016. p. 113--121.
+\newblock Available from: \url{http://doi.acm.org/10.1145/2960310.2960312}.
+
+\bibitem{alvidrez-weinstein}
+Alvidrez J, Weinstein RS.
+\newblock Early teacher perceptions and later student academic achievement.
+\newblock Journal of Educational Psychology. 1999;91(4):731--746.
+
+\bibitem{brophy}
+Brophy JE.
+\newblock Research on the self-fulfilling prophecy and teacher expectations.
+\newblock Journal of Educational Psychology. 1983;75(5):631--661.
+
+\bibitem{jussim-eccles}
+Jussim L, Eccles J.
+\newblock Social perception, social stereotypes, and teacher expectations:
+  Accuracy and the quest for the powerful self-fulfilling prophecy.
+\newblock Advances in Experimental Social Psychology. 1996;28:281--388.
+
+\bibitem{mazur-peer-instruction}
+Mazur E.
+\newblock Peer Instruction: A User's Manual.
+\newblock Prentice Hall; 1996.
+
+\bibitem{porter-multi}
+Porter L, Bailey~Lee C, Simon B, Cutts Q, Zingaro D.
+\newblock Experience Report: A Multi-classroom Report on the Value of Peer
+  Instruction.
+\newblock In: Proceedings of the 16th Annual Joint Conference on Innovation and
+  Technology in Computer Science Education; 2011. p. 138--142.
+
+\bibitem{porter-success}
+Porter L, Guzdial M, McDowell C, Simon B.
+\newblock Success in introductory programming: What works?
+\newblock Communications of the ACM. 2013;56(8).
+
+\bibitem{nrc-reaching}
+Council NR.
+\newblock Reaching Students: What Research Says About Effective Instruction in
+  Undergraduate Science and Engineering.
+\newblock Washington, DC: The National Academies Press; 2015.
+\newblock Available from:
+  \url{https://www.nap.edu/catalog/18687/reaching-students-what-research-says-about-effective-instruction-in-undergraduate}.
+
+\bibitem{rubin-live-coding}
+Rubin MJ.
+\newblock The Effectiveness of Live-coding to Teach Introductory Programming.
+\newblock In: Proceeding of the 44th ACM Technical Symposium on Computer
+  Science Education. SIGCSE '13. New York, NY, USA: ACM; 2013. p. 651--656.
+\newblock Available from: \url{http://doi.acm.org/10.1145/2445196.2445388}.
+
+\bibitem{barker-live-coding}
+Barker LJ, Garvin-Doxas K, Roberts E.
+\newblock What Can Computer Science Learn from a Fine Arts Approach to
+  Teaching?
+\newblock In: Proceedings of the 36th SIGCSE Technical Symposium on Computer
+  Science Education. SIGCSE '05. New York, NY, USA: ACM; 2005. p. 421--425.
+\newblock Available from: \url{http://doi.acm.org/10.1145/1047344.1047482}.
+
+\bibitem{willingham-book}
+Willingham DT.
+\newblock Why don't students like school?: A cognitive scientist answers
+  questions about how the mind works and what it means for the classroom.
+\newblock John Wiley \& Sons; 2009.
+
+\bibitem{crouch-demo}
+Crouch C, Fagen AP, Callan JP, Mazur E.
+\newblock Classroom demonstrations: Learning tools or entertainment?
+\newblock American Journal of Physics. 2004;72(6):835--838.
+\newblock doi:{10.1119/1.1707018}.
+
+\bibitem{miller-demo}
+Miller K, Lasry N, Chu K, Mazur E.
+\newblock Role of physics lecture demonstrations in conceptual learning.
+\newblock Phys Rev ST Phys Educ Res. 2013;9:020113.
+\newblock doi:{10.1103/PhysRevSTPER.9.020113}.
+
+\bibitem{hannay}
+Hannay JE, Dybå T, Arisholm E, , Sjøberg DIK.
+\newblock The effectiveness of pair programming: a meta-analysis.
+\newblock Information and Software Technology. 2009;51(7).
+
+\bibitem{mcdowell}
+McDowell C, Werner L, Bullock HE, Fernald J.
+\newblock Pair Programming Improves Student Retention, Confidence, and Program
+  Quality.
+\newblock Communications of the ACM. 2006;49(8):90--95.
+
+\bibitem{morrison-subgoals}
+Morrison BB, Margulieux LE, Guzdial M.
+\newblock Subgoals, Context, and Worked Examples in Learning Computing Problem
+  Solving.
+\newblock In: Proceedings of the Eleventh Annual International Conference on
+  International Computing Education Research. ICER '15. New York, NY, USA: ACM;
+  2015. p. 21--29.
+\newblock Available from: \url{http://doi.acm.org/10.1145/2787622.2787733}.
+
+\bibitem{morrison-parsons}
+Morrison BB, Margulieux LE, Ericson B, Guzdial M.
+\newblock Subgoals Help Students Solve Parsons Problems.
+\newblock In: Proceedings of the 47th ACM Technical Symposium on Computing
+  Science Education. SIGCSE '16. New York, NY, USA: ACM; 2016. p. 42--47.
+\newblock Available from: \url{http://doi.acm.org/10.1145/2839509.2844617}.
+
+\bibitem{margulieux}
+Margulieux LE, Guzdial M, Catrambone R.
+\newblock Subgoal-labeled Instructional Material Improves Performance and
+  Transfer in Learning to Develop Mobile Applications.
+\newblock In: Proceedings of the Ninth Annual International Conference on
+  International Computing Education Research; 2012. p. 71--78.
+
+\bibitem{gick-holyoak}
+Gick ML, Holyoak KJ.
+\newblock The cognitive basis of knowledge transfer.
+\newblock In: Cormier SM, Hagman JD, editors. Transfer of learning:
+  Contemporary research and applications. Academic Press; 1987.
+
+\bibitem{guzdial-media}
+Guzdial M.
+\newblock Exploring hypotheses about media computation.
+\newblock In: Proceedings of the Ninth Annual International ACM Conference on
+  International Computing Education Research. ICER '13; 2013. p. 19--26.
+
+\bibitem{bouvier-context}
+Bouvier D, Lovellette E, Matta J, Alshaigy B, Becker BA, Craig M, et~al.
+\newblock Novice Programmers and the Problem Description Effect.
+\newblock In: Proceedings of the 2016 ITiCSE Working Group Reports. ITiCSE '16.
+  New York, NY, USA: ACM; 2016. p. 103--118.
+\newblock Available from: \url{http://doi.acm.org/10.1145/3024906.3024912}.
+
+\bibitem{repenning}
+Repenning A.
+\newblock Moving beyond syntax: lessons from 20 years of blocks programing in
+  {AgentSheets}.
+\newblock Journal of Visual Languages and Sentient Systems. 2017;3.
+\newblock doi:{10.18293/VLSS2017-010}.
+
+\bibitem{parnin-expertise}
+Parnin C, Siegmund J, Peitek N.
+\newblock On the Nature of Programmer Expertise.
+\newblock In: Psychology of Programming Interest Group Workshop 2017; 2017.
+
+\bibitem{parsons}
+Parsons D, Haden P.
+\newblock Parsons Programming Puzzles: A Fun and Effective Learning Tool for
+  First Programming Courses.
+\newblock In: Proceedings of the 8th Australasian Conference on Computing
+  Education - Volume 52. ACE '06. Darlinghurst, Australia, Australia:
+  Australian Computer Society, Inc.; 2006. p. 157--163.
+\newblock Available from:
+  \url{http://dl.acm.org/citation.cfm?id=1151869.1151890}.
+
+\bibitem{ericson-ebook}
+Ericson B, Moore S, Morrison B, Guzdial M.
+\newblock Usability and Usage of Interactive Features in an Online Ebook for CS
+  Teachers.
+\newblock In: Proceedings of the Workshop in Primary and Secondary Computing
+  Education. WiPSCE '15. New York, NY, USA: ACM; 2015. p. 111--120.
+\newblock Available from: \url{http://doi.acm.org/10.1145/2818314.2818335}.
+
+\end{thebibliography}
 
